---
title: Comparing gating results
code_folding: fold
header-includes:
- \usepackage{bbm}
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=FALSE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=TRUE)
la('flowcy')
la('cruisedat')
library(popcycle)
library(tibble)
library(dplyr)
library(purrr)
repodir = "/home/shyun/repos/cruisedat"
outputdir = file.path(repodir, "export")
outputdir = "~/repos/cruisedat/export"
figdir = "~/Dropbox/research/usc/flow-cytometry/figures"
```

# Note of caution

This code is legacy code, mainly to load the "gates" (hard clustering
boundaries) from the original cruise's data, and to perform the gating.

You can obtain specific instructions for the datasets and for running this code
by contacting (emailing) the author Sangwon Hyun.

The R functions used here (e.g. `process_cruise_data()`) can be found in the
`sangwon-hyun/cruisedat` github repository (commit number
"916ad03ff36574ddaab6a79a3d533c772b708d82", also found here:
https://github.com/sangwon-hyun/cruisedat/commit/916ad03ff36574ddaab6a79a3d533c772b708d82
)

It also relies on *an earlier version of* the `popcycle` R package; use
`remotes::install_github("seaflow-uw/popcycle@035a9904fe4177e7a373d540b3217874ce3ee7d4")`
to install that version

# Obtain and save the gates information

Load data from database file `MGL1704.db`.
(This code is taken directly from `cruisedat::process_cruise_data()`.)

```{r, eval = FALSE}
## Synechococcus, Picoeukariotes are single gates, so simple matrices containing
## polygon vertices are returned.

## This step is important: the following code relies on an earlier version of
## popcycle
## remotes::install_github("seaflow-uw/popcycle@035a9904fe4177e7a373d540b3217874ce3ee7d4")
## library(popcycle)
poly_synecho <- get_synecho_gates()
save(poly, file = file.path(outputdir, "synecho-gates.Rdata"))
poly_picoeuk <- get_picoeuk_gates()
save(poly, file = file.path(outputdir, "picoeuk-gates.Rdata"))

## Prochlorococcus is more complicated, since the gating is done on a
## per-cytogram basis.
la('cruisedat')
res = get_prochloro_gates(mc.cores = 8)## plot=TRUE) ## This produces plots
list2env(res, envir = globalenv())
save(polylist, total_biomass_list, prochloro_biomass_list, total_count_list,
     prochloro_count_list, min.pe, file = file.path(outputdir, "prochloro-gates.Rdata"))
```

## Prochloro processing and plots

Now, plot the biomass and count **proportions**.

```{r}
load(file = file.path(outputdir, "prochloro-gates.Rdata"))

## 1. First, make 3-minute resolution prochlorococcus counts/biomass
prop_biomass = (prochloro_biomass_list %>% unlist) / (total_biomass_list %>% unlist)
names(prop_biomass) = names(prochloro_biomass_list)

prop_count = (prochloro_count_list %>% unlist) / (total_count_list %>% unlist)
names(prop_count) = names(prochloro_count_list)

## 2. Now, hourly aggregation of prochlorococcus counts/biomass.
times = names(prochloro_biomass_list)
obj = cruisedat::get_membership(times)
membership = obj$membership
time = obj$time

hourly_prochloro_biomass <- sapply(membership, function(memb){
  prochloro_biomass_list[memb] %>% unlist() %>% sum() })
hourly_total_biomass <- sapply(membership, function(memb){
  total_biomass_list[memb] %>% unlist() %>% sum()  })
names(hourly_prochloro_biomass) = names(hourly_total_biomass) = time
hourly_prop_biomass = hourly_prochloro_biomass / hourly_total_biomass

hourly_prochloro_count <- sapply(membership, function(memb){
  prochloro_count_list[memb] %>% unlist() %>% sum() })
hourly_total_count <- sapply(membership, function(memb){
  total_count_list[memb] %>% unlist() %>% sum()  })
names(hourly_prochloro_count) = names(hourly_total_count) = time
hourly_prop_count = hourly_prochloro_count / hourly_total_count

## 3. Make four plots, two 3-minutely, then two hourly plots.
pdf(file = file.path(figdir, paste0("prochloro-gating.pdf")),
    width = 12, height = 8)

par(mfrow = c(2,2))

plot(prop_biomass,
     ylab = "Biomass proportion", xlab = "", lty = 1, type = 'l',
     main = "(Prochloro biomass) / (total biomass), over time",
     axes = FALSE)
cruisedat::add_date_ticks_from_dates(names(prop_biomass))
abline(h=seq(from=0, to =1, by=0.1), col='grey80', lty=2)

plot(prop_count,
     ylab = "Count proportion", xlab = "", type = 'l',
     main = "(#Prochloro) / (#total), over time",
     axes  =  FALSE)
abline(h=seq(from=0, to =1, by=0.1), col='grey80', lty=2)
cruisedat::add_date_ticks_from_dates(names(prop_biomass))

plot(hourly_prop_biomass, 
     ylab = "Biomass proportion", xlab = "", lty = 1, type = 'l',
     main = "HOURLY\n(Prochloro biomass) / (total biomass), over time",
     axes = FALSE)
abline(h=seq(from=0, to =1, by=0.1), col='grey80', lty=2)
cruisedat::add_date_ticks_from_dates(time)

plot(hourly_prop_count, 
     ylab = "Count proportion", xlab = "", type = 'l',
     main = "HOURLY\n(#Prochloro) / (#total), over time",
     axes = FALSE)
cruisedat::add_date_ticks_from_dates(time)
abline(h=seq(from=0, to =1, by=0.1), col='grey80', lty=2)

graphics.off()
```

Now, obtain the prochloro biomass and count proportions from OUR 3d 10 cluster model.

```{r}
## ## Load 10 cluster 3d model
## cvres10 = flowcy::blockcv_summary(2, 64, 10, 7, nrep = 10)

## Reorder clusters in order of diam_mid()
## res <- flowcy::reorder_clust(cvres10$bestres)

datadir = "/home/sangwonh/repos/flowmix/paper-data" 
cvres_filename = paste0("cvres-", paste0(2,"-", 64,"-", 10), ".Rdata")
load(file = file.path(datadir, cvres_filename), verbose = TRUE)
res = cvres$bestres
class(res) = "flowmix" ## for back-compatibility
res$prob = res$pie ## for back-compatibility
res <- flowmix::reorder_clust(res)

## Make plot
pdf(file=file.path(figdir, "3d-10clust", "3d-prochloro-prob.pdf"),
    width=8,height=4)
par(mar=c(5.1, 4.1, 2.1, 2.1))
cols = RColorBrewer::brewer.pal(10, "Paired")
## cols = RColorBrewer::brewer.pal(10, "Set3")
iclusts = c(10)
flowmix::plot_prob(res, iclusts, main="", cols=cols)##, main = "HOURLY\n(Prochloro biomass) / (total biomass), over time")
lines(hourly_prop_biomass[-c(1:48)],### TODO: VERY AD HOC
      lty = 1, type = 'l',lwd=2,
      col='grey40')
     ## ylab = "Biomass proportion", xlab = "", ,
     ## main = "(Prochloro biomass) / (total biomass), over time",
     ## axes = FALSE)
dim(res$pie)
length(hourly_prop_biomass)
## abline(h=seq(from=0, to =1, by=0.1), col='grey80', lty=2)
legend("topleft",
       lwd = c(3, rep(3, length(iclusts))),##, 2),
       col=c('grey40', cols[iclusts]),##, "red"),
       lty = 1,
       legend=c("Gated Prochlorococcus\nbiomass proportion", paste0("Cluster ", iclusts, " probability")),
       bg="white")##, "Sum of clusters"))
graphics.off()
```

# Synecho and Picoeuks

Produce the gated results for synecho and picoeuks.

```{r}
## Obtain the gates
poly_synecho <- get_synecho_gates()
poly_picoeuk <- get_picoeuk_gates()

## Load |ylist|, |ybin_list|, |counts_list|, |biomass_list|, |X|
load(file = file.path(outputdir, "MGL1704-hourly.Rdata"), verbose = TRUE)

## Make two plots
relmat = cruisedat::gate_synecho_picoeuk(poly_synecho, ylist,
                                         ybin_list, counts_list, biomass_list, X)
cruisedat::plot_synecho_picoeuk(relmat,
                                res = res,
                                iclusts = 8,
                                plotname = "synecho-gating.pdf", celltype = "synecho")


relmat = cruisedat::gate_synecho_picoeuk(poly_picoeuk, ylist, ybin_list,
                                         counts_list, biomass_list, X, mc.cores=1)
cruisedat::plot_synecho_picoeuk(relmat,
                                res = res,
                                iclusts = c(2,5),
                                plotname = "picoeuk-gating.pdf", celltype = "picoeuk")


```

Also produce scatterplots of data with gates written on top!

```{r}
## Obtain the gates 
poly_synecho <- cruisedat::get_synecho_gates()
poly_picoeuk <- cruisedat::get_picoeuk_gates()

## Load |ylist|, |ybin_list|, |counts_list|, |biomass_list|
load(file = file.path(outputdir, "MGL1704-hourly.Rdata"), verbose = TRUE)

## Load 10 cluster 3d model
cvres10 = blockcv_summary(2, 64, 10, 7, nrep = 10)

## Reorder clusters in order of diam_mid()
res <- reorder_clust(cvres10$bestres)

## Setup
cex.fac = 1
ybin_list_diam_rescaled = ybin_list[-(1:12)]
biomass_list = biomass_list[-(1:12)]
TT = length(ybin_list_diam_rescaled)
times = names(ybin_list_diam_rescaled)

## Produce synecho and picoeuk gating plots.
start.time = Sys.time()
mclapply(1:TT, function(tt){
  printprogress(tt, TT, start.time = start.time)

  ## Rescale the diam_mid
  ybin_list_diam_rescaled[[tt]][,"diam_mid"] = ybin_list_diam_rescaled[[tt]][,"diam_mid"] %>%
    cruisedat::rescale_reshift_diam( hardcode=TRUE)

  ## Make plots
  png(file=file.path(figdir, "synecho-picoeuk", paste0("synecho-picoeuk-gate-boundary-", tt, ".png")),
                     width=1200, height=600)
  par(mfrow = c(1,2))
  dims = c(1,2)
  flowcy::one_dim_scatterplot(ybin_list_diam_rescaled, res, tt,
                              countslist = biomass_list,
                              dims = dims,
                              cex.fac = cex.fac)
  polygon(poly_picoeuk[,2],
          poly_picoeuk[,1],
          lwd=2)
  legend("topleft",
         legend="Picoeuk",
         bty='n',
         cex=2)
  title(main=times[tt],cex.main=2)

  dims = c(3,1)
  flowcy::one_dim_scatterplot(ybin_list_diam_rescaled, res, tt,
                              countslist = biomass_list,
                              dims = dims,
                              cex.fac = cex.fac)
  polygon(poly_synecho[,1],
          poly_synecho[,2],
          lwd=2)
  legend("topleft",
         legend="Synecho",
         bty='n',
         cex=2)
  title(main=times[tt],cex.main=2)
  graphics.off()
}, mc.cores = 8)
```

# Automatic detection from the existing data using ~flowDensity~

Prochlorococcus

```{r}
plot(res.tab[,dims],
     pch = 16,
     col = rgb(0, 0, 1, 0.1))
polygon(cruisedat::fsc_to_diam(poly[,1]),
        poly[,2],
        col = rgb(1,0,0,0.1),
        lwd = 3,
        border = "red")

## Get the gating parameters
gate = popcycle::get.gating.params.latest(dbfilename)

poly = gate$gates.log$prochloro$poly %>% log

names = c(gate$gates.log$prochloro$x,
          gate$gates.log$prochloro$y)
res.tab = cbind(log(opptable)[,c("fsc_small", "diam_mid", "chl_small", "pe")],
                opptable[,"Qc_mid"])
plot(res.tab[,names], type='p', pch=16, col=rgb(0,0,1,0.1))

abline(h=2,v=2)
abline(h=0.2,v=0.2)
polygon(poly[,1],
        poly[,2])

```

Some tidbits from the vignette:
- A gating strategy here means the **sequence of 2D** gates needed to apply
  **one at a time** on a FCS file to eventually extract the cell subset of
  interest. <--- This is probably because gating is done VISUALLY.
- Arguments to `flowDensity()`:
  - `percentile`: In the single cluster that is found using nonparametric
    density estimation, the 95\% region.
- From Francois' seaflow paper (https://www.nature.com/articles/s41597-019-0292-2)
  - "Prochlorococcus particles are clustered using a supervised clustering
    algorithm that emulates a sequential bivariate gating strategy based on cell
    density."
	- So, for Prochloro

The vignette here:
https://bioconductor.org/packages/devel/bioc/vignettes/flowDensity/inst/doc/flowDensity.html
was used. Section 3.3 is particularly useful.

This is interesting because our clustering is based on the biomass
representation of the data (i.e. each particle don't have equal weight, but
instead has a biomass weight that is proportional to `diam_mid^3`, in
clustering), but we can still produce counts from these gating regions.
