---
title: Subsampling for confidence intervals
author: Justin, Mattias, Francois, Jacob, 
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Download subsampled results

Here's a shell script to download all the summary files.

```{sh, eval=FALSE}
nsim=150
for isim in $(seq 1 1 $nsim) ; do
  mkdir /home/sangwonh/Dropbox/research/usc/hpc-output/subsample/subsample-"$isim"
  scp sangwonh@discovery.usc.edu:scratchdir/output/subsample/subsample-"$isim"/summary.RDS /home/sangwonh/Dropbox/research/usc/hpc-output/subsample/subsample-"$isim"/.
done
```

If I don't want to authorize 2FA 30 times, I can put in a little more work. On the server side, I can do something like this:
```{sh, eval=FALSE}
nsim=150
mkdir ~/scratchdir/output/subsample/subsample-summaries
for isim in $(seq 1 1 $nsim) ; do
	echo $isim
	cp ~/scratchdir/output/subsample/subsample-"$isim"/summary.RDS  ~/scratchdir/output/subsample/subsample-summaries/summary-"$isim".RDS
done
```
Then, download using one scp command.

```{sh, eval=FALSE}
mkdir /home/sangwonh/Dropbox/research/usc/hpc-output/subsample/subsample-summaries
from=sangwonh@discovery.usc.edu:scratchdir/output/subsample/subsample-summaries/*.RDS
to=/home/sangwonh/Dropbox/research/usc/hpc-output/subsample/subsample-summaries/.
scp $from $to
```

Bonus script: see progress in simulations so far

```{sh, eval=FALSE}
nsim=200
for isim in $(seq 175 1 $nsim) ; do
	echo $isim
	ls -f ~/scratchdir/output/subsample/subsample-"$isim"/ | wc -l
done
```

# KL divergence

Calculate the subsampled models' *reordered* cluster means, as follows/.

(The reordering is done by minimizing the KL divergence of the responsibilities,
of a subset of points.)

```{r}
## Load original particle-level data
obj = readRDS("~/repos/cruisedat/export/MGL1704-hourly-paper-1d-diam-not-binned.RDS")
ylist_particle = obj$ylist
X = dat$X


## Load original model
cvres = cv_summary(destin = file.path("~/Dropbox/research/usc/hpc-output/blockcv-2-76-5"))
origres = cvres$bestres
origres$prob = origres$pie
class(origres) = "flowmix"

## Compile the reordered cluster means
nsim = 150##200
TT = length(dat$ylist)
numclust = 5
mns = array(0, dim = c(TT, nsim, numclust))
source("subsample-helpers.R")

for(isim in 1:nsim){
  printprogress(isim, nsim)

  ## Load the CV results
  cvres = readRDS(file = file.path("~/Dropbox/research/usc/hpc-output/subsample/subsample-summaries",
                                   paste0("summary-", isim, ".RDS")))

  ## Gather the results for subsampling
  newres = predict(cvres$bestres, newx = X)
  class(newres) = "flowmix"

  ## Reorder the new res.
  source("subsample-helpers.R")
  newres = newres %>% reorder_kl(origres, ylist_particle, fac = 100, verbose=TRUE)

 ## Return the means
  mns[,isim,] = newres$mn[,1,]
}
saveRDS(mns, file=file.path("~/Desktop", "subsample-mns.Rdata"))
```

Then, produce the final plots:

```{r}
pdf(file = file.path("~/Desktop", "subsample-allmeans.pdf"),
    width = 25, height = 15)
par(mfrow = c(2, 3))
cols = RColorBrewer::brewer.pal(5, "Set2")[c(1,3,2,5,4)]
for(iclust in 1:5){
  plot_1d(ylist = dat$ylist,
          countslist = dat$countslist,
          cex_data = 3)
  matlines(mns[,,iclust], col = cols[iclust] %>% adjustcolor(alpha.f = 0.2),
           lty = 1, lwd = .1)
  matlines(origres$mn[,1,iclust], col = "black", lty = 1, lwd = 2.5)
  matlines(origres$mn[,1,iclust], col = "yellow", lty = 1, lwd = 2)
  mn_oneclust = mns0[,,iclust]
  for(s in 0.025){##c(0.025, 0.05, 0.1)){
    quantiles = apply(mn_oneclust, 1, quantile, probs=c(s, 1-s)) %>% t() 
    quantiles %>% matlines(lwd=.5,lty=1, col = "black")
  }
}
plot_1d(ylist = dat$ylist,
        countslist = dat$countslist,
        res = origres,
        cex_data = 3,
        reorder_clusters = TRUE)
graphics.off()
```


# Other code
```{r, eval=FALSE}
## Setup
la('flowmix')
source("subsample-helpers.R")

## Load data
dat = readRDS("~/repos/flowmix/paper-data/MGL1704-hourly-paper-1d-diam.RDS")

## Load the original data
obj = readRDS("~/repos/cruisedat/export/MGL1704-hourly-paper-1d-diam-not-binned.RDS")
ylist_particle = obj$ylist

## Load original results
cvres = cv_summary(destin = file.path("~/Dropbox/research/usc/hpc-output/blockcv-2-76-5"))
origres = cvres$bestres
origres$prob = origres$pie
class(origres) = "flowmix"

## We need to extract the column means
orig_mean_ord =  origres$mn[,1,] %>% colSums() %>% order(decreasing = TRUE)
origres = origres %>% reorder_clust(ord = orig_mean_ord)
assertthat::assert_that(all(origres$mn[,1,] %>% colSums() %>% order(decreasing = TRUE) == 1:5))

png(file = file.path("~/Desktop", paste0("subsampled-compare.png")), width=1200, height=3000)
par(mfrow=c(10, 3))
for(isim in 1:10){
  cvres = readRDS(file = file.path("~/Dropbox/research/usc/hpc-output/subsample/subsample-summaries",
                                   paste0("summary-", isim, ".RDS")))
  dat = readRDS("~/repos/flowmix/paper-data/MGL1704-hourly-paper-1d-diam.RDS")
  class(cvres$bestres) = "flowmix"
  newres = predict(cvres$bestres, newx = dat$X)
  class(newres) = "flowmix"
  
  ## Calculate new responsibilities
  resp_orig <- Estep(origres$mn,
                     origres$sigma,
                     origres$prob,
                     ylist = ylist_particle,
                     numclust = origres$numclust,
                     first_iter = TRUE)
  resp_new <- Estep(newres$mn,
                    newres$sigma,
                    newres$prob,
                    ylist = ylist_particle,
                    numclust = newres$numclust,
                    first_iter = TRUE)
  assertthat::assert_that(all(sapply(resp_orig, dim) == sapply(resp_new, dim)))

  ## Match using KL divergences
  matchres = kl_from_responsibilities(resp_new, resp_orig)
  kls = matchres$kls
  ordmat = matchres$ordmat

  ## Reorder orders using KL divergences
  best_ord = ordmat %>% .[which.min(kls),]
  newres_reordered_kl = newres %>% reorder_clust(ord = best_ord)

  ## Also reordered using average cluster means
  mean_ord =  newres$mn[,1,] %>% colSums() %>% order(decreasing=TRUE)
  newres_reordered_mn = newres %>% reorder_clust(ord = mean_ord)

  cols = plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = origres, reorder_clusters = FALSE)
  title(main = "Original", cex=2)

  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = newres_reordered_kl, reorder_clusters = FALSE)
  title(main = "Subsampled + Reordered\n (KL)")

  orig_mean_ord =  origres$mn[,1,] %>% colSums() %>% order(decreasing = TRUE)
  newres_reordered_mn = newres %>% reorder_clust(ord = mean_ord)

  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = newres_reordered_mn, reorder_clusters = FALSE)
  title(main = "Subsampled + Reordered\n (by avg of cluster means)", cex=2)

  rm(resp_orig)
  rm(resp_new)
}
graphics.off()

## png("~/Desktop/plot.png", width=1000, height=3000)
## par(mfrow = c(10, 2))
for(ii in 1:10){

  ## Reorder orders
  ords = order(kls, decreasing = FALSE)
  best_ord = ordmat %>% .[ords[ii],]
  newres_reordered = newres %>% reorder_clust(ord = best_ord)

  ## Make two plots
  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = newres_reordered, reorder_clusters = FALSE)
  title(main = "Subsampled + Reordered")
  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = origres, reorder_clusters = FALSE)
  title(main = "Original", cex=2)
  legend("topright", bty="n", legend = paste0("KL divergence = ", kls[ords[ii]] %>% round()), cex=2)

}


## saveRDS(resp_orig, file = file.path("~/Desktop", "resp_orig.RDS"))
## saveRDS(resp_new, file = file.path("~/Desktop", "resp_new.RDS"))
## resp_new = readRDS(file = file.path("~/Desktop",  "resp_new.RDS"))
## resp_orig = readRDS(file = file.path("~/Desktop",  "resp_orig.RDS"))

matchres = kl_from_responsibilities(resp_new, resp_orig)
kls = matchres$kls
ordmat = matchres$ordmat

## Plot the results
plot(kls, type='l')
abline(v=which.min(kls), lwd=3, col='red')
best_ord = ordmat %>% .[which.min(kls),]
best_ord = ordmat %>% .[which.max(kls),]

la('flowmix')
dat = readRDS("~/repos/flowmix/paper-data/MGL1704-hourly-paper-1d-diam.RDS")
class(newres) = class(origres) = "flowmix"
png("~/Desktop/plot.png", width=1000, height=3000)
par(mfrow = c(10, 2))
for(ii in 1:10){

  ## Reorder orders
  ords = order(kls, decreasing = FALSE)
  best_ord = ordmat %>% .[ords[ii],]
  newres_reordered = newres %>% reorder_clust(ord = best_ord)

  ## Make two plots
  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = newres_reordered, reorder_clusters = FALSE)
  title(main = "Subsampled + Reordered")
  plot_1d(ylist = dat$ylist, countslist = dat$countslist, res = origres, reorder_clusters = FALSE)
  title(main = "Original", cex=2)
  legend("topright", bty="n", legend = paste0("KL divergence = ", kls[ords[ii]] %>% round()), cex=2)
}
graphics.off()
```


# Make plots

Load data.

```{r}
library(flowmix)
dat = readRDS("~/repos/flowmix/paper-data/MGL1704-hourly-paper-1d-diam.RDS")
X = dat$X
```

Load subsampled results.

```{r}
la('flowmix')
newreslist = list()
nsim = 30
png(file = file.path("~/subsample", "subsample-grid.png"), width = 1000*3, height = 500*3)
par(mfrow = c(5, 6))
for(isim in 1:nsim){

  ## Load new model
  cvres = readRDS(file = file.path("~/Dropbox/research/usc/hpc-output/subsample/subsample-summaries",
                                   paste0("summary-", isim, ".RDS")))
  newres = predict(cvres$bestres, newx = X)
  class(newres) = "flowmix"
  plot_1d(res = newres,
          ylist = dat$ylist,
          countslist = dat$countslist,
          cex_dat = 1,
          time = dat$time)
  newreslist[[isim]] = newres
}
graphics.off()
 ```

Visualizing all of them together:

```{r}
## Load original model
cvres = cv_summary(destin = file.path("~/Dropbox/research/usc/hpc-output/blockcv-2-76-5"))

## Start a plot 
nsim = 100
plotfilename = file.path("~/subsample", paste0("model-all.png"))
png(file = plotfilename, width = 1000, height = 500)
plot_1d(res = NULL,
        ylist = dat$ylist,
        countslist = dat$countslist,
        cex_dat = 1,
        time_axis = TRUE,
        time = dat$time)

## Load a nsim
for(isim in 1:nsim){

  ## Reorder new models.
  newres = newreslist[[isim]] %>% reorder_clust()
  mns = newres$mn %>% .[,1,]
  prob = newres$prob
  numclust = newres$numclust

  ## Match clusters
  ## clusts = match_clust(mn_new = newres$mn,
  ##                      mn_orig = cvres$bestres$mn)


  cols = RColorBrewer::brewer.pal(numclust, "Set3")

  ## Add lines
  matlines(mns,
           type='l',
           pch = 15,
           ## col=rgb(1,0,0,0.2),
           col=cols,
           lty=1, lwd=2)
}
  matlines(cvres$bestres$mn %>% .[,1,],
           type='l',
           pch = 15,
           ## col=rgb(1,0,0,0.3),
           ## col=cols,
           col = "black",
           lty=1, lwd=3)
graphics.off()
```


Now, trying to match the clusters.

```{r}

cvres = cv_summary(destin = file.path("~/Dropbox/research/usc/hpc-output/blockcv-2-76-5"))



res = cvres$bestres
res$prob = res$pie ## for back-compatibility
```
