---
title: Producing figures of the 1D 5-cluster model
date: "`r Sys.Date()`"
code_folding: fold
header-includes:
- \usepackage{bbm}
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
author: Sangwon Hyun, Mattias Rolf Cape, Francois Ribalet, Jacob Bien
---

# Setup #

This script uses the flowmix package from github repository commit number
`206aa0575ea2df56f168f42dc335acf002aed458`
(https://github.com/sangwon-hyun/flowmix/commit/206aa0575ea2df56f168f42dc335acf002aed458).

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, warning=FALSE,
                      message=FALSE, cache=FALSE)
library(gridExtra)
library(tidyverse)
library(dplyr)
library(flowmix)

## This is where the output should (normally) go
figdir = "figures"

## This is where the data is
datadir = "../paper-data" 

## Temporarily load the latest flowmix package.
la('flowmix')
```
	
	
# 1d 5-cluster model results #

This section contains code to produce the plots for the 1d data 5-cluster
analysis in the paper (Figure 7, 12, 13, and Table 2).

First, run the simulations from `gradients2-run.R`, into a destination directory
`destin`. Then, use this code to summarize the results.

```{r, eval = FALSE, echo = FALSE}
cv_summary(destin = destin, nfold = 5, nrep = 10,
           save = TRUE, filename = "summary.RDS")##, subfolder="orig")
```

This will save the summary of the results into `file.path(destin, "summary.RDS")`.

In this script, I'm loading this file directly, from `./paper-data/1d-cvres.rds`

```{r load-model}
## Load the summary
## cvres = readRDS(file = file.path(destin, "summary.RDS"))

## Instead, load this summary
cvres_filename = "1d-cvres.rds"
cvres = readRDS(file = file.path(datadir, cvres_filename))

## For back-compatibility; prob was called pie
cvres$bestres$prob = cvres$bestres$pie 
class(cvres$bestres) = "flowmix"
```

We'll also load the 1d data from the directory `../paper-data`. (Normally, the
data is saved to `file.path(destin, "meta.Rdata")`)

```{r load-1d-data}
## load(file.path(cvres$destin, "meta.Rdata")) ## This is one way to load the data
dat = readRDS(file = file.path("../paper-data", "MGL1704-hourly-paper-1d-diam.RDS"))
countslist = dat$countslist
ylist = dat$ylist
```

## Figure 12 ##

Figure 12 is produced using this code:

```{r figure-12, fig.width = 12, height = 5}
## Plot CV scores
## pdf(file = file.path(figdir, "1d-cvscores.pdf"), width=12, height=5)
plot_cvscore(cvres$cvscore.mat)
## graphics.off()
```

## Figure 7 ##

Figure 7 is produced using this code:

```{r figure-7, fig.width = 12, fig.height = 5}
## Plot fitted mean.
## pdf(file = file.path(figdir, "1d-means.pdf"), width=12, height=5)
plot_1d(ylist = ylist, countslist = countslist,
        res = cvres$bestres,
        cex_data = 1) %>%  invisible()
mtext("Diameter", side = 2, line = 2.5, cex = 1.5)
## graphics.off()
```

## Table 2  ##

Table 2 can be produced using this code:

```{r table-2, results='asis'}
cvres$bestres$alpha %>% t() %>% round(3) %>% as.matrix() %>% knitr::kable() %>% print() ##%>%  xtable::xtable(digits=digits)
cvres$bestres$beta  %>% setNames(paste0("Cluster-", 1:5)) %>% do.call(cbind, .) %>% `colnames<-`(paste0("Cluster-", 1:5)) %>%
  round(3) %>% as.matrix() %>% knitr::kable() %>% print()##%>%  xtable::xtable(digits=digits)
```

## Figure 11 ##

Figure 11 can be produced using this code:

```{r covariates-plot, fig.height = 15, fig.width = 20}
mats = list(alpha = cvres$bestres$alpha, beta = cvres$bestres$beta)

## Plot covariates
allnames = unique(c(colnames(mats$alpha), rownames(mats$beta[[1]])))
allnames = allnames[-which(allnames == "intp")]
Xsmall = cvres$bestres$X %>% as_tibble %>% select(one_of(allnames)) %>% select(sort(names(.)))
inds = round(seq(from = 1, to = ncol(Xsmall), length=5))

##' Make ticks from rownames of res$X (only for 1d data).
##' @param res Object of class |flowmix|.
add_date_ticks <- function(res){
  dates = sapply(as.Date(rownames(res$X)), 
                 toString)

  nums = as.numeric(as.factor(dates))
  left_ticks = sapply(sort(unique(nums)),function(ii){min(which(nums==ii))})
  left_ticks = c(left_ticks, res$TT)
  mid_ticks = sapply(sort(unique(nums)),function(ii){mean(which(nums==ii))})
  dates_mid_ticks = dates[round(mid_ticks)]
  axis(1, at=left_ticks, labels=FALSE)
  axis(1, at=mid_ticks, labels = dates_mid_ticks, tick=FALSE, las=2)
  axis(2)
}


## pdf(file = file.path(figdir, "1d-covariates.pdf"), width=20, height=15)
par(mfrow=c(2,2))
for(ii in 1:4){
  ind = (inds[ii]+1):inds[ii+1]
  cols = RColorBrewer::brewer.pal(length(ind), "Set2")
  matplot(Xsmall, axes=FALSE, col='grey', type='l', lwd=.5, lty=1, ylab="",xlab="")
  add_date_ticks(cvres$bestres)
  matlines(Xsmall[,ind], col=cols, lwd=3, lty=1)
  legend("topright", col=cols, lwd=3, lty=1, legend = colnames(Xsmall)[ind])
}
## graphics.off()
```

## Figure 13 ##

Figure 13 is produced using the second block (starting with `# A 5 x 5 version of this`):

```{r figure-13, eval=FALSE}
load(file.path(cvres$destin, "meta.Rdata")) ## load ylist
## png(file=file.path(figdir, "1d-all-models.png"), width=5000, height=4000)
cv_gridsize = 10
par(mfrow = c(cv_gridsize, cv_gridsize))
for(ialpha in 1:cv_gridsize){
  for(ibeta in 1:cv_gridsize){
    res = cvres$bestreslist[[paste0(ialpha, "-", ibeta)]]
    res$prob = res$pie ## For back-compatibility; prob was called pie
    plot_1d(ylist = ylist, countslist = countslist, res = res,
            cex_data = 1)
    title(main = c(paste0("in-sample: ", round(min(res$objective),3)), "   ",
                   paste0("CV score : ", round(cvres$cvscore.mat[ialpha,ibeta],3))),
          cex.main=3)
    if(all(c(ialpha, ibeta) == cvres$min.ind)) box(lwd=10,col='blue')
  }
}
## graphics.off()

## A 5 x 5 version of this.
load(file.path(cvres$destin, "meta.Rdata")) ## load ylist
## png(file=file.path(figdir, "1d-all-models-reduced.png"), width=5000/2, height=3000/2)
cv_gridsize = 10
par(mfrow = c(cv_gridsize/2, cv_gridsize/2 ))
par(mar=c(0,0,5,0))
for(ialpha in c(1,3,5,7,10)){
  for(ibeta in c(1,4,6,8,10)){
    print(c(ialpha, ibeta))
    res = cvres$bestreslist[[paste0(ialpha, "-", ibeta)]]
    res$prob = res$pie ## For back-compatibility; prob was called pie
    plot_1d(ylist = ylist, countslist = countslist, res = res,
            no_axis=TRUE, cex_clust_label=NULL, omit_label=TRUE,
            omit_band=TRUE,
            cex_data = 1.7)
    title(main = c(paste0("In-sample: ", round(min(res$objective),3)), "   ",
                   paste0("CV score : ", round(cvres$cvscore.mat[ialpha,ibeta],3))),
          cex.main=4)##3.5)
    if(all(c(ialpha, ibeta) == cvres$min.ind)) box(lwd=10,col='blue')
  }
}
## graphics.off()
```


# 3d 10-cluster model results #


```{r load-3d-model}
### Load the model 
cvres_filename = paste0("cvres-", paste0(2,"-", 64,"-", 10), ".Rdata")

### cvres = blockcv_summary(2, 64, 10, 7, nrep = 5) 

### save(cvres, file = file.path(datadir, cvres_filename)) 

load(file = file.path(datadir, cvres_filename), verbose = FALSE)
res = cvres$bestres
destin = cvres$destin

## Quick adjustments, for back-compatibility
class(res) = "flowmix" ## for back-compatibility
res$prob = res$pie ## for back-compatibility
```

## Table 3,4,5,6 ##

Table 3,4,5,6 are produced using this code:

```{r, results = 'asis'}
for(iclust in 1:10){
  onebeta = res$beta[[iclust]]
  colnames(onebeta) = c("diam", "red", "orange")
  knitr::kable(onebeta, digits=10, caption = paste0("Beta coefficientns for Cluster ", iclust)) %>% print()
}

t(res$alpha) %>%  knitr::kable(digits=10, caption = paste0("Alpha coefficients")) %>% print()
```

## Figure 8 ##

Figure 8 is produced using this code:

```{r figure-8, fig.width = 12, fig.height=5}
## Rescale diameters (not used anymore since diam in the data is in original
## range, and is shifted/scaled to be in the same scale as CHL and PE only at 
## simulation time.)
source("gradients2-helpers.R")

## Temporary *undoing* of the scaling in the estimated model.
range.diam = c(-0.6706111,  1.3070264)
range.chl =  c(0.101313, 7.955767)
range.pe = c(0.101313, 7.955767)
  
## Modify (rescale/shift) the res$mn
width.diam = range.diam[2] - range.diam[1]
idim = 1
res$mn = res$mn[,idim,, drop=FALSE] / max(range.chl) * width.diam + min(range.diam)

## Undoing the scaling in res$sigma
res$sigma = cvres$bestres$sigma[,idim, idim, drop=FALSE] 
myscale = max(range.chl) * width.diam
res$sigma = res$sigma / myscale

## Load the data
datobj = readRDS(file.path(datadir, "MGL1704-hourly-paper-1d-diam.RDS"))
ylist = datobj$ylist
countslist = datobj$countslist

## Make the 1d data+model plot
res$dimdat = 1
cols = plot_1d(ylist = ylist, countslist = countslist, res, scale=TRUE, omit_band=TRUE, omit_label=TRUE)

## Additionally, add y axis label
mtext("Diameter", side=2, line=2.5)

## Additionally, write the cluster numbers with the same colors
res <- reorder_clust(res)
TT = length(ylist)
for(iclust in 1:res$numclust){
  y = res$mn[1, 1, iclust]
  if(iclust==3) y = y + 0.05
  if(iclust==4) y = y - 0.05
  text(x = res$TT/50 - 5, y = y,
       label = paste0("Cluster ", iclust), cex = 1, col="black")##col=cols[iclust])
}
```

It's worth explaining a bit about data rescaling for Gradients 2. The three
dimensions of the original cytogram data (e.g.  `MGL1704-hourly-paper.Rdata` or
in `MGL1704-hourly.Rdata`) all have a range roughly between 0 and 8, as you can see
here:

```{r figure-8-1, results = 'asis'}
datobj = readRDS(file = file.path(datadir, "MGL1704-hourly-paper.RDS"))
datobj$ylist[[2]] %>% apply(2, range) %>% knitr::kable() %>% print()
```

This is the result of shifting and stretching the diameter dimension (the first
of the three dimensions), which originally had a range of:

```{r figure-8-2, results='asis'}
datobj = readRDS(file.path(datadir, "MGL1704-hourly-paper-1d-diam.RDS"));
datobj$ylist[[1]] %>% range() %>% cbind() %>% `colnames<-`(c("diam_mid")) %>% knitr::kable() %>% print()
```

<!-- |----------| -->
<!-- | -0.6706111| -->
<!-- |  1.3070264| -->
<!-- |----------| -->

to be the same range as the other two dimensions (Chl and PE), between $0$ and
$7.955767$. So the code above is just reversing this shifting and scaling.

## Figure 9 ##

Figure 9 is produced using this code:

```{r figure-9, fig.width = 24, fig.height = 6}
plot_3d_fourpanels <- function(obj, ## Understandably, data (ylist) might not be in the object.
                               ylist, countslist = NULL, ## The time point of interest, out of 1:TT
                               tt, ## Other options.
                               ## 2d scatterplot options
                               show.xb.constraint = FALSE, cex.fac.2d = 1, par_cex_2d = 1,
                               pt_col = rgb(0 ,0, 1, 0.1), ## 3d scatterplot options
                               cex.fac.3d = 1, ## 3d scatterplot options
                               destin = NULL){

  ## Define layout
  par(mfrow = c(1, 4))
  ## par(oma = c(3, 1, 2, 1)) ## setting outer margin

  ## Setup
  TT = length(ylist)
  assertthat::assert_that(tt %in% 1:TT)
  all.y = do.call(rbind, ylist)
  only_plot_cytograms = (is.null(obj))
  if(!only_plot_cytograms){
    obj = reorder_clust(obj)
    mns = obj$mn
    numclust = obj$numclust
    p = ncol(obj$X)
  }

  ## Scale the biomass (|countslist|) by the total biomass in that cytogram.
  counts_sum = sapply(countslist, sum)
  fac = median(counts_sum)
  countslist = lapply(countslist, function(counts)counts/sum(counts) * fac)


  ###############################
  ## Make the three data plots ##
  ###############################
  ## par(mar=c(1,1,3,1))
  par(mar = c(5.1, 5.1, 4.1, 2.1))
  dimslist = list(1:2, 2:3, c(3,1))
  for(dims in dimslist){
    ## one_dim_scatterplot(ylist, obj, tt,
    scatterplot_2d(ylist, obj, tt,
                   countslist = countslist,
                   dims = dims,
                   cex_fac = cex.fac.2d,
                   pt_col = pt_col,
                   lwd = 2)
  }
  par(cex=0.8)

  par(mar=c(1,1,3,1))
  ## phis = c(10,50)
  phis = c(10)
  for(phi in phis){
    one_3d_plot(ylist, obj, tt, countslist = countslist, phi = phi,
                cex.fac = cex.fac.3d)
  }
}

## Load 3d model
cvres_filename = paste0("cvres-", paste0(2,"-", 64,"-", 10), ".Rdata")
load(file = file.path(datadir, cvres_filename), verbose = FALSE)
res = cvres$bestres
class(res) = "flowmix" ## for back-compatibility
res$prob = res$pie ## for back-compatibility

datobj = readRDS(file = file.path(datadir, "MGL1704-hourly-paper.RDS"))
ylist = datobj$ylist
countslist = datobj$countslist
tt = 10
plot_3d_fourpanels(res, ylist, countslist, tt = tt, destin = destin,
                   cex.fac = 10, pt_col = rgb(0,0,1,0.2), cex.fac.3d = 0.4)
```

## Figure 10 ##

Figure 10 is produced using this code:

(The file `prochloro-gates.Rdata` comes from `paper-code/gating.Rmd`).

```{r figure-10-load-data}
## Load the results from traditional gating of Prochlorococcus
load(file = file.path(datadir, "prochloro-gates.Rdata"))

## 1. First, make 3-minute resolution prochlorococcus counts/biomass
prop_biomass = (prochloro_biomass_list %>% unlist) / (total_biomass_list %>% unlist)
names(prop_biomass) = names(prochloro_biomass_list)

prop_count = (prochloro_count_list %>% unlist) / (total_count_list %>% unlist)
names(prop_count) = names(prochloro_count_list)

## 2. Now, hourly aggregation of prochlorococcus counts/biomass.
times = names(prochloro_biomass_list)
obj = get_membership(times) ## borrowed from the "cruisedat" package
membership = obj$membership
time = obj$time

hourly_prochloro_biomass <- sapply(membership, function(memb){
  prochloro_biomass_list[memb] %>% unlist() %>% sum() })
hourly_total_biomass <- sapply(membership, function(memb){
  total_biomass_list[memb] %>% unlist() %>% sum()  })
names(hourly_prochloro_biomass) = names(hourly_total_biomass) = time
hourly_prop_biomass = hourly_prochloro_biomass / hourly_total_biomass
```


```{r figure-10, fig.width = 8, fig.height = 4}
## Make plot 
## pdf(file = file.path(figdir, "3d-prochloro-prob.pdf"),
##     width = 8,height = 4)
par(mar = c(5.1, 4.1, 2.1, 2.1))
cols = RColorBrewer::brewer.pal(10, "Paired")
iclusts = c(10)
flowmix::plot_prob(res, iclusts, main = "", cols = cols)
lines(hourly_prop_biomass[-c(1:48)],
      lty = 1, type = 'l',lwd = 2,
      col='grey40')
legend("topleft",
       lwd = c(3, rep(3, length(iclusts))),
       col=c('grey40', cols[iclusts]),
       lty = 1,
       legend=c("Gated Prochlorococcus\nbiomass proportion", paste0("Cluster ", iclusts, " probability")),
       bg="white")
## graphics.off()
```



## Figure 14 ##

Figure 14 is produced using this code:

```{r figure-14, fig.width = 30, fig.height = 15}
## ## Load 3d model
## cvres_filename = paste0("cvres-", paste0(2,"-", 64,"-", 10), ".Rdata")
## load(file = file.path(datadir, cvres_filename), verbose = FALSE)
## res = cvres$bestres
## class(res) = "flowmix" ## for back-compatibility
## res$prob = res$pie ## for back-compatibility

## ## Load 3d data
## datobj = readRDS(file = file.path(datadir, "MGL1704-hourly-paper.RDS"))
## ylist = datobj$ylist
## countslist = datobj$countslist

tt = 10 ## WHAT is this?
plot_3d(res, ylist, countslist, tt = tt, cex.fac = 10, pt_col = rgb(0,0,1,0.2),
        cex.fac.3d = 0.4,
        lat = datobj$lat,
        lon = datobj$lon)
```


<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height:0;"></div>

