---
title: 1D 5-cluster data example in the paper
code_folding: fold
header-includes:
- \usepackage{bbm}
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
author: Sangwon Hyun, Mattias Rolf Cape, Francois Ribalet, Jacob Bien
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=TRUE)
library(gridExtra)
library(tidyverse)
library(dplyr)
library(flowmix)

## This is where the output should go
figdir = "~/repos/flowmix/paper-code/figures"

## This is where the parent directory of the output
outputdir = "/home/sangwonh/Dropbox/research/usc/hpc-output"

## This is specific directory of the output
destin = file.path(outputdir, "blockcv-2-76-5")
```


This script contains code to produce the plots for the 1d data 5-cluster
analysis in the paper (Figure 7, 12, 13, and Table 2).

First, run the simulations from `gradients2-run.R`, into a destination directory
`destin`. Then, use this code to summarize the results.

```{r, eval=FALSE, echo=FALSE}
cv_summary(destin = destin, nfold = 5, nrep = 10,
           save = TRUE, filename = "summary.RDS")##, subfolder="orig")
```

The summarized results are in `file.path(destin, "summary.RDS")`.

(In the next code block, I've replaced destin with the `./paper-code`)

```{r}
## Load the summary
cvres = readRDS(file = file.path(destin, "summary.RDS"))

## For back-compatibility; prob was called pie
cvres$bestres$prob = cvres$bestres$pie 
```

None of the `flowmix` objects contain the data (by design) so we need to load
the data separately.

The most accurate way is to load from `file.path(destin, "Meta.Rdata")`.

In this script, we'll load from the directory `../paper-data`.

```{r}
## load(file.path(cvres$destin, "meta.Rdata")) ## This is one way to load the data
dat = readRDS(file = file.path("../paper-data", "MGL1704-hourly-paper-1d-diam.RDS"))
countslist = dat$countslist
ylist = dat$ylist
```


Figure 12 is produced using this code:

```{r}
## Plot CV scores
pdf(file = file.path(figdir, "1d-cvscores.pdf"), width=12, height=5)
plot_cvscore(cvres$cvscore.mat)
graphics.off()
```

Figure 7 is produced using this code:

```{r}
## Plot fitted mean.
pdf(file = file.path(figdir, "1d-means.pdf"), width=12, height=5)
plot_1d(ylist = ylist, countslist = countslist,
        res = cvres$bestres,
        cex_data = 1)
mtext("Diameter", side = 2, line = 2.5, cex = 1.5)
graphics.off()
```

Table 2 can be produced using this:

```{r, eval=FALSE}
cvres$bestres$alpha %>% t() %>% signif(3) %>% as.matrix() ##%>%  xtable::xtable(digits=digits)
cvres$bestres$beta %>% t() %>% signif(3) %>% as.matrix() ##%>%  xtable::xtable(digits=digits)
```

(Not in paper) Covariates can be plotted like this:

```{r, eval=FALSE}
mats = list(alpha = cvres$bestres$alpha, beta = cvres$bestres$beta)

## Some investigations
cvres$bestres$X %>% as_data_frame %>%
  select(p1, p2, p3, contains("phosphate"), contains("density"), contains("ugos")) %>%
  cbind(ind =1:nrow(cvres$bestres$X)) -> Xsmall
reshape2::melt(Xsmall, id = "ind") -> Xsmall.long
ggplot2::qplot(x=ind, y=value, color=variable, data=Xsmall.long, geom="line")

## Plot covariates
allnames = unique(c(rownames(mats$alphamat), rownames(mats$betamat)))
Xsmall = cvres$bestres$X %>% as_tibble %>% select(one_of(allnames)) %>% select(sort(names(.)))
inds = round(seq(from = 1, to = ncol(Xsmall), length=5))

pdf(file = file.path(figdir, "1d-covariates.pdf"), width=20, height=15)
par(mfrow=c(2,2))
for(ii in 1:4){
  ind = (inds[ii]+1):inds[ii+1]
  cols = RColorBrewer::brewer.pal(length(ind), "Set2")
  matplot(Xsmall, axes=FALSE, col='grey', type='l', lwd=.5, lty=1, ylab="",xlab="")
  add_date_ticks(cvres$bestres)
  matlines(Xsmall[,ind], col=cols, lwd=3, lty=1)
  legend("topright", col=cols, lwd=3, lty=1, legend = colnames(Xsmall)[ind])
}
graphics.off()
```


Figure 13 is produced using the second block (starting with `# A 5 x 5 version of this`):

```{r, eval=FALSE}
load(file.path(cvres$destin, "meta.Rdata")) ## load ylist
png(file=file.path(figdir, "1d-all-models.png"), width=5000, height=4000)
cv_gridsize = 10
par(mfrow = c(cv_gridsize, cv_gridsize))
for(ialpha in 1:cv_gridsize){
  for(ibeta in 1:cv_gridsize){
    res = cvres$bestreslist[[paste0(ialpha, "-", ibeta)]]
    res$prob = res$pie ## For back-compatibility; prob was called pie
    plot_1d(ylist = ylist, countslist = countslist, res = res,
            cex_data = 1)
    title(main = c(paste0("in-sample: ", round(min(res$objective),3)), "   ",
                   paste0("CV score : ", round(cvres$cvscore.mat[ialpha,ibeta],3))),
          cex.main=3)
    if(all(c(ialpha, ibeta) == cvres$min.ind)) box(lwd=10,col='blue')
  }
}
graphics.off()


## A 5 x 5 version of this.
load(file.path(cvres$destin, "meta.Rdata")) ## load ylist
png(file=file.path(figdir, "1d-all-models-reduced.png"), width=5000/2, height=3000/2)
cv_gridsize = 10
par(mfrow = c(cv_gridsize/2, cv_gridsize/2 ))
par(mar=c(0,0,5,0))
for(ialpha in c(1,3,5,7,10)){
  for(ibeta in c(1,4,6,8,10)){
    print(c(ialpha, ibeta))
    res = cvres$bestreslist[[paste0(ialpha, "-", ibeta)]]
    res$prob = res$pie ## For back-compatibility; prob was called pie
    plot_1d(ylist = ylist, countslist = countslist, res = res,
            no_axis=TRUE, cex_clust_label=NULL, omit_label=TRUE,
            omit_band=TRUE,
            cex_data = 1.7)
    title(main = c(paste0("In-sample: ", round(min(res$objective),3)), "   ",
                   paste0("CV score : ", round(cvres$cvscore.mat[ialpha,ibeta],3))),
          cex.main=4)##3.5)
    if(all(c(ialpha, ibeta) == cvres$min.ind)) box(lwd=10,col='blue')
  }
}
graphics.off()
```


<!-- # Analysis -->

<!-- * **Cluster 5** The usual culprits are all there -- lagged sunlight ("par") and -->
<!--   sea surface temperature and salinity all affect multiple clusters. In -->
<!--   particular, the lagged sunlight ("p1/p2/p3/p4") plays a large role in cluster -->
<!--   5, which is visually apparent, and neat that our model captures this. -->
<!-- * **Cluster 4** Biochemical factors like PP (Net primary productivity of Carbon -->
<!--   per unit volume), nitrate, and $O_2$, as well as wind/current like -->
<!--   "ugos/vgos/ftle" seems to play a large role in cluster 4. -->
<!-- * **Cluster 2** Chlorophyll (and Phosphate) plays a large role in cluster 2's -->
<!--   means and probabilities -- this is clear if you look at CHL (and to a lesser -->
<!--   extent, "phosphate_WOA_clim") over time, it is high on June 1-4, and -->
<!--   fluctuates similarly to the prevalence of the cluster during that dates.. -->
<!-- * **Cluster 1** wind/current like "ugos/vgos/northword_wind/sla" seems to play a -->
<!--   large role in cluster 1. -->
