---
title: 1D simulation example in the paper (varying the number of clusters)
code_folding: fold
header-includes:
- \usepackage{bbm}
author: Sangwon Hyun
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=FALSE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=TRUE)
la('flowcy')
library('flexmix')
library(Matrix)
figdir = "./figures"
datadir = "./data"
```

# Note of caution

This code is legacy code that accompanies the `paper-code/simulations.Rmd`
script, and runs the experiments for examining how mis-specifying the number of
clusters affects the out-of-sample model fit.


# Summary

Produces two pdf files `noisy-covariate-fit.pdf` and `noisy-covariate-power.pdf`, for the 
figure label `\label{fig:numclusts-results}`.


## Simulation steps

Internally, each job is numbered by data types 80-89, which is relevant for
`blockcv-generate-data.R`.

All of the following scripts are contained in the zip file `all-scripts.R`

Scripts to run on a SLURM server:

1. **Generate the data** once, using the script: `./blockcv-numclust-gendat.R` 

3. **Run the simulations** `./run-sims-numclust.sh`
  - which calls `run-blockcv-2-9-x.slurm` over datatypes 9, numclust = 2, ..., 10.
  - which calls `blockcv-numclust.R` which runs the simulation. (this script is
    different from `blockcv.R`)

4. **Summarize the simulations on server** `./run-summarize-numclust.sh`
  - which calls `run-summarize-numclust.slurm`
  - which calls `summarize-numclust.R` on data types 9, numclust = 2, ..., 10.
  - which calls `blockcv-summarize-sim()`.

5. **Download results on my computer**, `./run-summarize-client-numclust.sh`
  - which downloads
  - `scp hpc:/scratch/sangwonh/output/blockcv-2-9-[ ]/summary/.` to
   `/Dropbox/research/usc/hpc-output/blockcv-2-9-[ ]/summary/.`for [ ] = 2, ..., 10.


# Code

Visualizing the data once.

```{r viz-numclust-sim}
obj = generate_data_1d_pseudoreal_from_cv(datadir=datadir##"~/repos/cruisedat/export",
                                          bin = TRUE, plot = TRUE, dat.gridsize=50)
```

Now, we download the data once. The steps are similar to before.
- Make some summaries directly on the server, using `summarize-numclust.sh`.
- Download the server, using `run-summarize-client.sh` to download these results.

```{r}
numclusts = c(2,3,4,5,6,7,8)
nsim = 30
probs = c(0.05, 0.95, 0.25, 0.75)

## Empty CV score objects
avg_cvscores = rep(NA, length(numclusts))
quantile_cvscores = matrix(NA, nrow=length(numclusts), ncol=4)
all_cvscores = matrix(NA, nrow=length(numclusts), ncol=nsim)

## Empty out-of-sample score objects
avg_oos_scores = rep(NA,  length(numclusts)) ## new
quantile_oos_scores = matrix(NA, nrow=length(numclusts), ncol=4)
all_oos_scores = matrix(NA, nrow=length(numclusts), ncol=nsim)

## Generate LARGE new dataset (large = large n_t$) from new model
datobj = generate_data_1d_pseudoreal_from_cv(datadir="~/repos/cruisedat/export", nt=3000)

blocktype = 2
datatype = 9
for(iclust in 1:length(numclusts)){
  printprogress(iclust, 1:length(numclusts), "iclusts", fill=TRUE)
  numclust = numclusts[iclust]
  ## datadir = "~/stagedir/output"
  destin = file.path(datadir,
                     paste0("blockcv-", blocktype, "-", datatype, "-", numclust))
  load(file = file.path(destin, "summary", "bestreslist.Rdata"))

  nsim = length(bestreslist)
  ## New: calculate the REAL CV score 
  for(isim in 1:nsim){
    printprogress(isim, nsim, "sim", fill=TRUE)
    bestres = bestreslist[[isim]]
    if(is.null(bestres))next
    all_oos_scores[iclust, isim] = objective_newdat(ylist = datobj$ylist,
                                                    countslist = datobj$countslist,
                                                    res  =  bestres)

  }
}

## Calculate the average and quantiles
for(iclust in 1:length(numclusts)){
  numclust = numclusts[iclust]
  avg_oos_scores[iclust] = mean(all_oos_scores[iclust,], na.rm=TRUE)
  quantile_oos_scores[iclust,] = quantile(all_oos_scores[iclust,], probs = probs,
                                      na.rm=TRUE) ## middle 90% quantile.
}

## Labeling quantile CV scores
rownames(quantile_cvscores) = numclusts
colnames(quantile_cvscores) = probs
rownames(all_cvscores) = numclusts
names(avg_cvscores) = numclusts

## Labeling out of sample scores
rownames(quantile_oos_scores) = numclusts
colnames(quantile_oos_scores) = probs
rownames(all_oos_scores) = numclusts
names(avg_oos_scores) = numclusts

pdf(file=file.path(figdir, "numclusts-cvscores-new.pdf"), width=5, height=5)
par(mar=c(5.1,4.1,3.1,2.1))
all_numclusts = 2:8
plot(NA,
     xlim = c(2, 8),
     ylim = range(all_oos_scores, na.rm=TRUE),
     ylab = "Out-of-sample NLL",
     xlab = "Number of clusters used in estimation (K)",
     axes=FALSE)
abline(v=5, col='orange', lwd=3)
abline(h=seq(from=0.40, to=0.48, by=0.02), col='grey80', lty=2)
## text(x=5, y=0.47, label="True underlying K")
legend('topright', col=c(rgb(0,0,0,0.2), 'black', 'orange'),
       lwd = c(NA, 3,3),
       pch = c(16, NA,NA),
       legend = c("One simulation", "Average", "True K"),
       bg = "white")
axis(1, at=1:10);
axis(2)
pch = 1:30
for(iclust in 1:length(numclusts)){
  numclust = numclusts[iclust]
  points(y = all_oos_scores[iclust, ],
         x = rep(numclust, ncol(all_oos_scores)),
         ## type = 'p',
         col = rgb(0,0,0,0.2),
         cex = 1,
         ## pch = sapply(1:30, toString)
         pch = 16
         )
}
lines(y = avg_oos_scores,
      x = numclusts,
      lwd = 3, type = 'l')
graphics.off()

```
