---
title: 1D simulation example in the paper (noisy covariates)
code_folding: fold
header-includes:
- \usepackage{bbm}
author: Sangwon Hyun
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=FALSE, warning=FALSE,
                      message=FALSE, eval=TRUE, cache=TRUE)
library(flexmix)
library(Matrix)
load_all('~/repos/flowmix/flowmix') ## Replace
figdir = "./figures"
```

# Note of caution

This code is legacy code that accompanies the `paper-code/simulations.Rmd`
script, and runs the experiments for examining how adding noise to the
covariates affects the model performance.

# Summary

This script produces three plots:

- `noisy-covariates-data.pdf`, and `noisy-covariates.pdf` for the figure label
  `\label{fig:noisy-covariates-data}`.
- `noisy-covariate-fit.pdf` for the figure label
  `\label{fig:noisy-covariates-results}`.


## Simulation steps

Internally, each job is numbered by data types 80-89, which is relevant for
`blockcv-generate-data.R`.

All of the following scripts are contained in the zip file `all-scripts.R`

Scripts to run on a SLURM server:

1. Generate the data once: `./run-sim-gendat.sh` 
  - which calls `run-sim-gendat.slurm`.
  - which calls `blockcv-sim-gendat.R`

2. Get the maximum regularization parameters: `./run-sim-maxres.sh` 
  - which calls `run-sim-maxres.slurm`.
  - which calls `blockcv-sim-maxres.R`

3. **Run the simulations** `./run-sims.sh`
  - which calls `run-sim.slurm` over datatypes 80-89
  - which calls `blockcv-sim.R` which runs the simulation. (this script is
    different from `blockcv.R`)

4. **Summarize the simulations on server** `./run-sim-summarize.sh`
  - which calls `run-summarize.slurm`
  - which calls `summarize.R` on data types 80-89.
  - which calls `blockcv-summarize-sim()`.

5. **Download results on my computer**, `./run-summarize-client.sh`
  - which downloads
  - `scp hpc:/scratch/sangwonh/output/blockcv-2-80-2/summary/.` to
   `/Dropbox/research/usc/hpc-output/blockcv-2-80-2/summary/.`

## Code for summarizing results

First, we want to visualize the **out-of-sample likelihoods** as a function of
$\sigma_{\text{add}}$.

```{r cv-error-viz}

##' From simulation results, loads |bestreslist| containing the chosen models
##' from CV, then calculates the out-of-sample log likelihood of these models
##' with respect to a large new dataset (cytograms) generated from the same
##' mechanism.
##'
##' @param blocktype CV block type
##' @param datatypes 80:89
##' @param numclust 2
##' @param cv_gridsize 7
##' @param nrep 5
##' @param outputdir Output directory on client computer
##' 
##' @return List of OOS scores.
outsample_loglik <- function(blocktype = 2, 
                             datatypes = c(80:89),
                             numclust = 2,
                             cv_gridsize = 7,
                             nrep = 5,
                             outputdir = "~/Dropbox/research/usc/hpc-output"
                             ){

  ## Generate large data once
  datobj_large = generate_data_1d_pseudoreal(datadir = datadir,##"~/repos/cruisedat/export",
                                             nt1 = 20000,
                                             beta_par = 0.3,
                                             p = 10)

  ## Calculate the out-of-sample scores
  ## datatypes = c(80:89)
  ## datatypes = c(80, 83, 86, 89)
  avg_oos = rep(NA, length(datatypes))
  quantile_oos = matrix(NA, nrow=length(datatypes), ncol = 4)
  nsim = 100
  all_oos = matrix(NA, nrow = length(datatypes), ncol = nsim)
  probs = c(0.05, 0.95, 0.25, 0.75)
  for(ii in 1:length(datatypes)){
    print(ii)
  
    ## Load the best model from each noise level
    datatype = datatypes[ii]
    destin = file.path(outputdir, paste0("blockcv-", blocktype, "-", datatype, "-", numclust))
    load(file = file.path(destin, "summary", "bestreslist.Rdata"))
    nsim = length(bestreslist)
    print(nsim)
    for(isim in 1:nsim){
      ## lapply(1:9, function(isim){
      bestres = bestreslist[[isim]]
      if(is.null(bestres)) next
      oos_loglik = objective_newdat(ylist = datobj_large$ylist,
                                   countslist = datobj_large$countslist,
                                   res = bestres)
      all_oos[ii, isim] = oos_loglik
    }
  }

  ## Obtain the mean & quantiles
  avg_oos = apply(all_oos, 1, mean, na.rm = TRUE) ##mean(all_oos[ii,])
  quantile_oos = apply(all_oos, 1, function(myrow){
    quantile(myrow, probs = probs, na.rm=TRUE)
  })

  ## Return results
  return(list(all_oos = all_oos,
              avg_oos = avg_oos,
              quantile_oos = quantile_oos))
} 


## Make plots 
datatypes = c(80:89)
sigmalist = seq(from=0,to=2.7, by=0.3)
lwd = 3
ooslist =  outsample_loglik(2, datatypes, 2)
list2env(ooslist, envir=.GlobalEnv)
pdf(file=file.path(figdir, "noisy-covariate-fit.pdf"), width=5, height=5)
par(mar = c(4.1, 4.1, 1.1, 2.1))
par(mgp = c(2.5, 1, 0))
plot(NA,
     ylab = "Out-of-sample NLL",
     xlab = expression(sigma["add"]),
     ylim = range(all_oos, na.rm=TRUE),
     ## ylim = range(c(the_best_loglik - 0.005, all_oos), na.rm=TRUE),
     xlim = range(sigmalist),
     axes = FALSE)
abline(h = seq(from=1.6, to=1.7, by = 0.01), col ='grey80', lty=2)
## abline(h = the_best_loglik, lwd=3, col='orange')
## text(y = the_best_loglik * 1.003, x=2, labels="Best attainable \n Neg. log likelihood (NLL)")
matlines(y = all_oos,
        x = sigmalist,
        type='p', pch=16, col=rgb(0,0,0,0.1))
axis(2); axis(1, at=sigmalist)##, las=3)##[c(1,2*(1:5))]);
lines(y = avg_oos,
      x = sigmalist,
      lwd=3, type='l')
legend("bottomright",
       col = c(rgb(0,0,0,0.1), "black"),
       lwd = c(NA, lwd),
       pch = c(16, NA),
       legend = c("NLL (Neg. log-likelihood)", "Average NLL"),
       bty="o",
       border="white",
       bg="white")
graphics.off()
```

## Visualizing the sparsity.

This figure is not used in the paper.

```{r viz-coef}
datatypes = c(80:89)
sigmalist = seq(from=0,to=2.7, by=0.3)[datatypes-79]

## Calculate power
thresh = 0
res = get_power(datatypes = datatypes,
                sigmalist = sigmalist,
                thresh = thresh) %>% select(-contains("alpha")) ##%>% select(-beta_par_prob)

## Make the plot
figdir = "~/Dropbox/research/usc/flow-cytometry/figures"
pdf(file = file.path(figdir, "noisy-covariate-power.pdf"), width=5, height=5)
plot_power(res)
graphics.off()

##' Plot power.
plot_power <- function(res){ 

  res %>% select(-datatypes) -> res
  res %>% select(sigma_add) %>%unlist -> sigmalist
  res %>% select(-sigma_add) -> res
  ## par(mar=c(3,3,3,3))
  par(mar = c(4.1, 4.1, 1.1, 2.1))
  par(mgp = c(2.5, 1, 0))
  matplot(NA, ylim = c(0,1), xlim = range(sigmalist),
          ylab = expression(P(hat(beta)!=0)),
          xlab = expression(sigma["add"]),
          axes = FALSE)
  matlines(y = res %>% select(contains("beta_par")),
           x = sigmalist,
           type = 'l', col = c(1,1),
           lwd = 3, lty = c(1,2))
  matlines(y = res %>% select(contains("beta_spurious_prob1")), x = sigmalist,
           type = 'l', lwd = 1, lty = 1, col=rgb(1,0,0,0.2))
  matlines(y = res %>% select(contains("beta_spurious_prob1")) %>% rowMeans, x = sigmalist,
           type = 'l', lwd = 3, lty = 1, col='red')
  matlines(y = res %>% select(contains("beta_spurious_prob2")), x = sigmalist,
           type = 'l', lwd = 1, lty = 1, col=rgb(1,0,0,0.2))
  matlines(y = res %>% select(contains("beta_spurious_prob2")) %>% rowMeans, x = sigmalist,
           type = 'l', lwd = 3, lty = 2, col='red')
  axis(2, at=seq(from=0,to=1, by=0.2)); axis(1, at=sigmalist)##, las=3)##[c(1,2*(1:5))]);
  legend(x=1.1, y=0.92,##"topright",
         lwd = c(3,3,3,3,1),
         lty = c(1,2,1,2,1),
         col=c("black","black","red","red",rgb(1,0,0,0.2)),
         legend=c("Sunlight, Clust 1",
             "Sunlight, Clust 2",
             "Avg Spurious, Clust 1",
             "Avg Spurious, Clust 2",
             "Each Spurious"), bty='n')
  return()
}

##' Obtain P(nonzero), and return as a data frame whose rows are for different
##' additive noise, and the columns are different probabilities; probabilities
##' for coefficients of each cluster separately, and together.
##'
##' @param datatypes 
##'
get_power <- function(blocktype = 2, 
                      datatypes = c(80:89),
                      sigmalist,
                      numclust = 2,
                      cv_gridsize = 7,
                      nrep = 5,
                      thresh = 0,
                      outputdir = "~/Dropbox/research/usc/hpc-output"){

  stopifnot(length(sigmalist) == length(datatypes))

  nsim_max=100
  beta_par_prob1 = beta_par_prob2 = c()
  beta_spurious_prob1 = beta_spurious_prob2 = matrix(NA, nrow = length(datatypes), ncol = 8)
  for(ii in c(1:length(datatypes))){ 
    print(ii)
    
    datatype = datatypes[ii]
    destin = file.path(outputdir,
                       paste0("blockcv-", blocktype, "-", datatype, "-", numclust))
    
    ## Obtain the best quantile CV scores
    load(file = file.path(destin, "summary", "bestreslist.Rdata")) ## This would normally work!
    nsim = length(bestreslist)

    ## Obtain the beta coefficients from simulations
    betas = lapply(bestreslist, function(bestres){
      if(!is.null(bestres)){
      beta = do.call(cbind, bestres$beta)
      ## order the two according to the intercept
      beta = beta[,order(beta[1,])]
      } else { return(NULL)}
    })
    betas = betas[which(sapply(betas, length) > 0)] ## temporary handling
  
    ## Count the interesting stuff (sunlight betas)
    beta_par_prob1[ii] = sapply(betas, function(beta)abs(beta["par",1]) > thresh) %>% mean
    beta_par_prob2[ii] = sapply(betas, function(beta)abs(beta["par",2]) > thresh) %>% mean

    ## Count the interesting stuff (spurious betas)
    spurious_nonzero1 = sapply(betas, function(beta){
      apply(beta[4:nrow(beta),], 1, function(myrow){
      (abs(myrow[1])>thresh)})
    })
    spurious_nonzero2 = sapply(betas, function(beta){
      apply(beta[4:nrow(beta),], 1, function(myrow){
      (abs(myrow[2])>thresh)})
    })
    beta_spurious_prob1[ii,] = apply(spurious_nonzero1, 1, mean)
    beta_spurious_prob2[ii,] = apply(spurious_nonzero2, 1, mean)
  }
  res = data.frame(datatypes = datatypes,
                   beta_par_prob1 = beta_par_prob1,
                   beta_par_prob2 = beta_par_prob2,
                   beta_spurious_prob1 = beta_spurious_prob1,
                   beta_spurious_prob2 = beta_spurious_prob2,
                   sigma_add = sigmalist)
  return(res)
}
```

## Visualizing the data

Here are the generating coefficients.

```{r viz-coef-table}
## Export the true covariates
obj = generate_data_1d_pseudoreal(nt1 = 200, beta_par = 0.3, p = 10)
list2env(obj, envir = globalenv())
print(xtable::xtable(cbind(alpha, beta)) )
```

Visualizing the data once.

```{r viz-data}
ymat = do.call(rbind, lapply(ylist, as.numeric))
pdf(file=file.path(figdir, "noisy-covariates-data.pdf"), width=7, height=5)
par(mar = c(5.1, 4.1, 1.1, 2.1))
par(mgp = c(2.5, 1, 0))
cex.fac = 1.3
par(cex.lab=cex.fac)
matplot(NA, xlim=c(0,100), ylim=c(-5, 7), axes=FALSE, ylab="Data", xlab="Time")
abline(h=seq(from=-30, to=30, by=1), lwd=.5, col='grey90', lty=2)
matpoints(ymat, type='p', pch=16, col=rgb(0,0,0,0.05), cex=0.8)
axis(1, cex.axis=cex.fac); axis(2, tick=TRUE, label=TRUE, cex.axis=cex.fac)
cols = RColorBrewer::brewer.pal( 3, "Set2")
cols = sapply(cols, function(col) adjustcolor( col, alpha.f = 0.5))
matlines(mnmat, type='l', lwd=.5, col=cols, lty=2)
points(mnmat[,1], type='p', lwd=2, col=cols[1], pch=15, cex=pie[,1]*3)
points(mnmat[,2], type='p', lwd=2, col=cols[2], pch=15, cex=pie[,2]*3)
legend("topleft", col=cols, lwd=2, pch=16, lty=1,
       legend=c(expression(Cluster~1~Mean~~mu["1t"]),
                expression(Cluster~2~Mean~~mu["2t"])),
       bg = "white",
       cex=cex.fac)
graphics.off()
```

Visualizing the noisy sunlight covariate:

```{r viz-covariates}
pdf(file.path(figdir, "noisy-covariates.pdf"), width=7, height=5) 
par(mar = c(5.1, 4.1, 1.1, 2.1))
par(mgp = c(2.5, 1, 0))
cex.fac = 1.3
par(cex.lab = cex.fac)
lwd = 2
cols = RColorBrewer::brewer.pal(4, "Set2")
matplot(NA, ylab="Sunlight covariate", xlab="Time", axes=FALSE,ylim=c(-5, 10), xlim = c(0,100))
abline(h=seq(from = -5, to = 10), lwd=.5, col='grey90', lty=3)
set.seed(0)
lines(obj$X[,1] + rnorm(nrow(X), 0, 2.7), type='l', lwd=lwd, lty=1,  col=cols[3],)
lines(obj$X[,1] + rnorm(nrow(X), 0, 1), type='l', lwd=lwd, lty=1,  col=cols[2],)
lines(obj$X[,1] + rnorm(nrow(X), 0, 0), type='l', lwd=5, lty=1,  col=1)##cols[1],)
lines(obj$X[,2], type='l', lwd=5, lty=1,  col=2)
lines(obj$X[,1] + rnorm(nrow(X), 0, 0.3), type='l', lwd=lwd, lty=1,  col=cols[1],)
axis(1)
axis(2)
legend("topright", col=c("red","black", cols[1:3]), lwd=c(5,5,rep(lwd,3)), lty=1, legend=c(expression(Changepoint~X[2]),
                                                                                   expression(Sunlight~X[1]),
                                                                                   expression(Noisy~Sunlight~sigma["add"]==0.3),
                                                                                   expression(Noisy~Sunlight~sigma["add"]==1),
                                                                                   expression(Noisy~Sunlight~sigma["add"]==2.7)),
       bg="white",
       cex = cex.fac)
legend("topright", col=c(""))
graphics.off()
```
