---
title: Estimating stability of coefficient estimates
author: Sangwon Hyun, Mattias Rolf Cape, Francois Ribalet, Jacob Bien
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# Setup

This script uses the flowmix package from github repository commit number
`206aa0575ea2df56f168f42dc335acf002aed458`
(https://github.com/sangwon-hyun/flowmix/commit/206aa0575ea2df56f168f42dc335acf002aed458).

```{r setup, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE,
                      cache=FALSE, warning=FALSE, message=FALSE,
                      cache.lazy = FALSE)
library(flowmix)
library(tidyverse)
library(xtable)
## source("~/repos/flowmix/graveyard.R")
```

# Code to conduct the subsampling bootstrap

Loading the data files required:

* Particle level data: `MGL1704-hourly-paper-1d-diam-not-binned.RDS`
  + This is non-binned cytogram data used for matching clusters.

* Summaries from estimated models in subsampled datasets:
  + `subsample-random-time/subsample-b-150/subsample-summaries`
  + Subsampling is done by randomly taking time blocks of length $T/2 = 150$
    (out of $T=296$ timepoints in total).
  + The results are in 100 files named like
    `subsample-random-time/subsample-b-150/sim-10` (sim-1 through sim-100).
  + These files are produced by running `run-subsample.slurm`, which calls
      `subsample.R` repeatedly on a high-performance computing environment with
      a slurm task management system.
	
* The original model is in `paper-data/1d-cvres.RDS`. (Internally, the
  experiment name is `blockcv-2-76-5`)

The following code is used to obtain all the subsampled models, in the object
`all-models.RDS`. 

```{r load-model}
## Load original model
## cvres = cv_summary(destin = file.path("~/Dropbox/research/usc/hpc-output/blockcv-2-76-5"))
datadir = "~/repos/flowmix/paper-data/"
cvres = readRDS(file.path(datadir, "1d-cvres.rds"))
origres = cvres$bestres
origres$prob = origres$pie
class(origres) = "flowmix"
origres = origres %>% reorder_clust()
```

This code was used to further summarize the results, but is currently not run in
this script.

```{r get-subsampled-models, eval = FALSE}
## Load original particle-level data (for cluster matching)
obj = readRDS("~/repos/cruisedat/export/MGL1704-hourly-paper-1d-diam-not-binned.RDS")
ylist_particle = obj$ylist
X = obj$X
rm(obj)

## Get all the models
nsim = 100
outputdir = "~/Dropbox/research/usc/hpc-output/subsample-random-time/subsample-b-150/subsample-summaries"
all_models = get_simulated_models(nsim, outputdir, origres, ylist_particle, X) 
## saveRDS(all_models, file = file.path(datadir, "all-models.RDS"))
```

## Table 7 and 8 ##

Assuming the code above has been run, the model summaries are loaded directly
from an RDS file `all-models.RDS`.

```{r load-summaries}
all_models = readRDS(file = file.path(datadir, "subsampling", "all-models.RDS"))
numclust = 5
```

Then, we calculate the beta coefficients' stability estimates.

```{r calculate-beta-stability, results='asis'}
## Make the frequency matrix
beta_list = lapply(all_models, function(a)a$beta)
beta_nonzero_list = beta_list %>% lapply(., function(beta){ lapply(beta, function(b) b==0) })
beta_nonzero = lapply(1:numclust, function(iclust){
  beta_nonzero_list %>% lapply(., function(b)b[[iclust]]) %>% Reduce("+", .)
}) %>% do.call(cbind, .)
nsim = 100
betamat_orig = beta_nonzero ##%>% t()
betamat_orig = nsim - betamat_orig
betamat_orig = betamat_orig / nsim

## ## Showing together with parameter estimates
## combined_bmat = lapply(1:numclust, function(iclust){
##   a1 = origres$beta %>% .[[iclust]] %>% .[-1,] %>% sapply(., function(cc){ if(cc==0) -Inf else cc})
##   a2 = betamat_orig  %>% .[-1,] %>% .[,iclust] %>% sapply(., function(cc){ if(cc==0) -Inf else cc})
##   cbind(coef=a1, freq=a2)
## }) %>% do.call(cbind, .)
## ## paste0("Cluster-", rep(1:5, each = 2)) %>% paste0(collapse = " & ")
## colnames(combined_bmat) = paste0("Cluster-", rep(1:5, each = 2))
## combined_bmat %>% knitr::kable() %>% print() ## %>% Matrix() %>% round(3) %>% print() %>% to.
```

Tables 7 and 8 (left panel) are produced using this code:

```{r table-7-8, results='asis'}
## Sorting within each cluster
numclust = 5
beta_ranked = lapply(1:numclust, function(iclust){

  ## Get the estimates and the stability
  a1 = origres$beta %>% .[[iclust]] %>% .[-1,] %>% sapply(., function(cc){ if(cc==0) -Inf else cc})
  a2 = betamat_orig %>% .[-1,iclust] %>% sapply(., function(cc){ if(cc==0) -Inf else cc})
  a = cbind(a1, a2)

  ## Reorder this by the stability
  a = a[order(a2, decreasing=TRUE),,drop=FALSE] %>% round(3)
  rn = rownames(a)

  ## Reformat
  a = a %>%
    as_tibble(rownames = NA) %>%
    rownames_to_column( var = "rowname") %>%
    mutate(a1 = sprintf("%.3f", a1),
           a2 = sprintf("%.2f", a2)) ##%>% as.matrix()

  ## convert the rownames to have \texttt{}
  texttt <- function(vec_of_strings){
    sapply(vec_of_strings, function(str) paste0("texttt{", str, "}"))
  }
  remove_clutter <- function(vec_of_strings){
    vec_of_strings = sapply(vec_of_strings, function(str) sub("_WOA_clim", "", str))
    vec_of_strings = sapply(vec_of_strings, function(str) gsub("surface_downward_northward_stress", "sdns", str))
    vec_of_strings = sapply(vec_of_strings, function(str) gsub("surface_downward_eastward_stress", "sdns", str))
    return(vec_of_strings)
  }
  ## a = a %>% mutate(rowname = texttt(rowname))
  a = a %>% mutate(rowname = (rowname))
  a = a %>% mutate(rowname = remove_clutter(rowname)) ## Not used now.

  ## Change back -Inf (which means zero) to cdot
  a = a %>%
    mutate(across(c(a1,a2), .fns = function(a) sub("-Inf", "$.$", a,  fixed = TRUE)))  %>% 
    rbind(c("Variable", "Estim.", "P(nonzero)"), .)
  colnames(a) = rep(paste0("Cluster ", iclust), 3)
  return(a)
})
beta_ranked[1:3] %>% do.call(cbind, .)  %>% knitr::kable() %>% print()
beta_ranked[4:5] %>% do.call(cbind, .)  %>% knitr::kable() %>% print()
```


<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height:0;"></div>
