---
title: Experiments for examining time-dependence
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
header-includes: 
  - \usepackage{bm} 
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

\newcommand{\one}{\mathbbm{1}}
\newcommand{\R}{\mathbb{R}}


# Note of caution

This code is legacy code that accompanies the `paper-code/time-dependence.Rmd`
script, and runs the experiments for examining time-dependence in the residuals.

You can obtain specific instructions for the datasets and for running this code
by contacting (emailing) the author Sangwon Hyun.

The R functions used here (e.g. `process_cruise_data()`) can be found in the
`sangwon-hyun/cruisedat` github repository (commit number
"916ad03ff36574ddaab6a79a3d533c772b708d82", also found here:
https://github.com/sangwon-hyun/cruisedat/commit/916ad03ff36574ddaab6a79a3d533c772b708d82
)

It also relies on *an earlier version of* the `popcycle` R package; use

```{r global_options, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=5, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE)
                      ## cache.lazy = FALSE)

## Load packages
library(tidyverse)
library(knitr)
library(here)
library(ggplot2)
library(MTS)
library(purrr)
library(plotly)
library(ggpmisc)
source("time-dependence-experiments-helpers.R")
la("flowmix")


## Output directory
## outputdir = "~/repos/flowmixapp/data/02-estimate/MGL1704/summaries/"
outputdir = "/home/sangwonh/repos/flowmix/paper-code/aoas-revision/3d-residuals"
maindir = "/home/sangwonh/repos/flowmix/paper-code/aoas-revision"
## outputdir = "~/scratchdir/output/flowmixapp/data/02-estimate/MGL1704/summaries/overtrim" ## on the server
```
 
# Video of 3d model 

# {.tabset}

```{r labeled-video, results='asis'}
for(ii in 1:2){

  if(ii == 1)cat("## Unlabeled", fill = TRUE)
  if(ii == 2)cat("## Labeled", fill = TRUE)
  
  ## Make video
  if(ii==1)videofile = file.path(maindir, "MGL1704-model-unlabeled.mp4") 
  if(ii==2)videofile = file.path(maindir, "MGL1704-model.mp4") 
  
  ## Embed video
  cat(paste0('<video width="1280" height="720" controls>  <source src="', 
             videofile,
             '" type="video/mp4"> </video>'))
  cat('\n\n')
}
```

# Preparing model and residuals

## Load 3d MGL1704 model

Load model from the new experiments.

```{r load-model-and-data}
modeldir = "~/repos/flowmixapp/data/02-estimate/MGL1704/summaries"
summary = readRDS(file =  file.path(modeldir, "summary-orig.RDS"))
origres = summary$bestres

datadir = "~/repos/flowmixapp/data/01-cytograms/MGL1704/data"
particle_datobj = readRDS(file =  file.path(datadir, "MGL1704-particle-datobj.RDS"))
times = particle_datobj$time
```

## Load the particle level data

```{r plot-particle-data, fig.width = 5, fig.height = 5}
datobj_2d = collapse_3d_to_2d(
    y = particle_datobj %>% .$ylist %>% .[[1]],
    counts = particle_datobj %>% .$countslist %>% .[[1]],
    dims = c(1,2)) %>% as_tibble()
time = particle_datobj %>% .$ylist %>% names() %>% .[[1]]
datobj_2d %>% 
  ggplot() +
  geom_point(aes(x = diam, y = chl, size = counts), col = rgb(0,0,1,0.01)) +
  scale_size_area() +
  ggtitle(paste0("Particle cytogram ", time))
```


## Do the coin-flip experiment

Doing the coin-flip experiment to draw particle memebership using
responsibilities.

(This is done on the server, in `scripts/flowmix-revision/bin.R`.)

```{r coin-flip, eval = FALSE}
## Particle data to use
ylist_particle = particle_datobj$ylist
countslist_particle = particle_datobj$countslist
ylist_particle = ylist_particle %>% lapply(as.matrix)

## Draw membership
resp <- Estep(origres$mn, origres$sigma, origres$prob, ylist = ylist_particle,
              numclust = origres$numclust, first_iter = TRUE)
set.seed(0)
drawslist = draw_membership(resp)

rm(particle_datobj)
rm(resp)

save(drawslist, file = file.path(outputdir, "drawslist-3d.Rdata"))


## Get the residuals
residuals_obj = get_residuals(ylist = ylist_particle,
                              countslist = countslist_particle,
                              res = origres,
                              drawslist = drawslist,
                              mc.cores = 1)

## Save it
save(residuals_obj, file = file.path(outputdir, "residuals-3d.Rdata"))
```

Now, binning the residuals once.

(This is also done on the server, in `scripts/flowmix-revision/bin.R`.)


```{r bin-residuals, eval = FALSE}
## Read the data
load(file = file.path(outputdir, "residuals-3d.Rdata"), verbose = TRUE)


## Reverse order of nested list (from time->clust to clust->time)
numclust = 10
residuals = residuals_obj$residuals_by_cluster
countslist = residuals_obj$countslist_by_cluster
resids_by_clust = cluster_then_time(residuals)
counts_by_clust = cluster_then_time(countslist)

## Make grid for binning
ylist_particle = particle_datobj$ylist
dat_grid = flowmix::make_grid(ylist_particle, gridsize = 100)
TT = length(resid_by_clust[[1]])
residuals = residuals %>% lapply(function(a){
  a = a %>% lapply(as_tibble) %>% bind_rows()
})
## dat_grid = flowmix::make_grid(residuals, gridsize = 40)
saveRDS(dat_grid, file = file.path(outputdir, "dat_grid.RDS"))


## Bin once
dat_grid = readRDS(file = file.path(outputdir, "dat_grid.RDS"))
binobj = list()
for(iclust in 1:numclust){
  print(iclust)
  obj = flowmix::bin_many_cytograms(ylist = resid_by_clust[[iclust]],
                                    qclist = counts_by_clust[[iclust]],
                                    dat_grid, mc.cores = 8, verbose = TRUE)
  binobj[[iclust]] = obj
}


## Save the binned residuals
saveRDS(binobj, file = file.path(outputdir, "binned-scaled-residuals-3d.RDS"))
```


Lastly, here's a command to move things from the server to a local directory.


```{sh download-from-server, eval = FALSE}
rsync discovery:scratchdir/output/flowmixapp/data/02-estimate/MGL1704/3d-residuals/* ~/repos/flowmix/paper-code/aoas-revision/3d-residuals/.
```


## Scaled residuals

Now, form **scaled** residuals.

(Scaling here means deconvolving; if the $k$'th cluster's covariance matrix was
$\Sigma_k \in R^{3 \times 3}$, then the residuals $r \in R^{3 \times 1}$ are
transformed by a left-multiplication by the inverse square root
$\Sigma^{-1/2}$.)

(Why are there fractional amounts in each bin?)

```{r scaled-residuals, fig.width = 15, fig.height = 11}
## Read the residual data in
load(file = file.path(outputdir, "residuals-3d.Rdata"), verbose = FALSE)

## ## Temporary
## tt = 1
## iclust = 1
## cbind(residuals_obj %>% .$residuals_by_cluster %>% .[[tt]] %>% .[[iclust]],
##       counts = residuals_obj %>% .$countslist_by_cluster %>% .[[tt]] %>% .[[iclust]]) %>% head()
## ## End of temporary: why are these all 

## Scale (deconvolve) the particle-level residuals
mu = origres$mn
prob = origres$prob
sigma = origres$sigma
numclust = 10
TT = dim(mu)[1]
all_scaled_resids = lapply(1:TT, get_scaled_resids_realdata_3d,
                           residuals_obj, sigma) 

## Okay, let's see if this deconvolving makes sense in one time point and one cluser
tt = 10
iclust = 3
allnames = c("diam", "chl", "pe")

## Before de-convolving
plist1 = lapply(list(1:2, 2:3, c(3,1)), function(inds){
  name1 = allnames[inds[1]]
  name2 = allnames[inds[2]]
  g2 = residuals_obj %>% .$residuals_by_cluster %>% .[[tt]] %>% .[[iclust]] %>% as_tibble() %>% 
    ggplot() +
    geom_point(aes(x = !!sym(name1), y = !!sym(name2)), alpha = .5) +
    labs(title = "Synecho residuals",
         subtitle = times[[tt]])
}) 

## After de-convolving
plist2 = lapply(list(1:2, 2:3, c(3,1)), function(inds){
  name1 = allnames[inds[1]]
  name2 = allnames[inds[2]]
  a = get_scaled_resids_realdata_3d(tt, residuals_obj, sigma)
  dat = a$resids %>% .[[iclust]] %>% as_tibble() %>% setNames(c("diam", "chl", "pe"))
  g1 = dat  %>% ggplot() +
    geom_point(aes(x = !!sym(name1), y = !!sym(name2)), alpha = .5) +
    labs(title = "Synecho residuals, deconvolved",
         subtitle = times[[tt]])
})

do.call(gridExtra::grid.arrange, c(c(plist1, plist2), ncol=3))
```

Bin these, for later plotting (done on the server) in `scripts/flowmix-revision/bin.R`.

```{r bin-scaled-residuals, eval = FALSE}
## Get the residuals
TT = nrow(prob)
numclust = 10
all_scaled_resids = lapply(1:TT, get_scaled_resids_realdata_3d,
                           residuals_obj, sigma)

## Reorganize data by cluster
numclust = 10
resid_by_clust = list()
counts_by_clust = list()
TT = length(residuals)
numclust = length(residuals[[1]])
for(iclust in 1:numclust){
  resids_thistime = list()
  counts_thistime = list()
  for(tt in 1:TT){
    resids_thistime[[tt]] = all_scaled_resids %>% .[[tt]] %>% .$resids %>% .[[iclust]]
    counts_thistime[[tt]] = all_scaled_resids %>% .[[tt]] %>% .$counts %>% .[[iclust]]
  }
  resid_by_clust[[iclust]]  = resids_thistime
  counts_by_clust[[iclust]]  = counts_thistime
}

## Make the grid
TT = length(residuals)
residuals = lapply(1:TT, function(tt){
  a = all_scaled_resids %>% .[[tt]]  %>% .$resids %>% lapply(as.data.frame) %>% bind_rows()
})
dat_grid = flowmix::make_grid(residuals, gridsize = 100)

## Bin the scaled residuals
binobj = list()
for(iclust in 1:numclust){
  print(iclust)
  obj = flowmix::bin_many_cytograms(ylist = resid_by_clust[[iclust]],
                                    qclist = counts_by_clust[[iclust]],
                                    dat_grid, mc.cores = 8, verbose = TRUE)
  binobj[[iclust]] = obj
}

## Save the result
saveRDS(binobj, file = file.path(outputdir, "binned-scaled-residuals-3d.RDS"))
```

# Visualize the (scaled) residuals

Before de-convolving.

```{r viz-2d-residuals, fig.width = 10, fig.height=20}
binobj = readRDS(file = file.path(outputdir, "binned-residuals-3d.RDS"))

## Load the dates
modeldir = "~/repos/flowmixapp/data/01-cytograms/MGL1704/data"
times = readRDS(file = file.path(modeldir, "MGL1704-datobj.RDS"))$time

## ## Saving them once.
## datadir = "~/repos/flowmix/paper-data"
## saveRDS(times, file=file.path(datadir, "binned-residuals-time.RDS"))

## Make 2d plots of a few snapshot in time
iclust = 3
glist = list()
dats_2d = lapply(1:9, function(ii){
  tt = ii * 20
  datobj_2d = collapse_3d_to_2d(
      y = binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]],
      counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]],
      dims = c(1,2)) %>%
    as_tibble() %>%
    add_column(time = times[tt])
}) %>% bind_rows()##.id = "date")
dats_2d = dats_2d %>% mutate(time = as.factor(time))

name1 = "diam"
name2 = "chl"
dats_2d %>% ggplot() +
  facet_wrap(~time) +
  geom_raster(aes(x = !!sym(name1), y = !!sym(name2), fill = counts)) +
  scale_fill_gradientn(colours = c("white", "blue"), guide="colorbar")  +
  coord_fixed() +
  ggtitle("Original residuals")
  ## theme_minimal() 
```


A few 2d (diam vs chl) scaled residual plots.

```{r viz-2d-scaled-residuals, fig.width = 10, fig.height=17}
binobj = readRDS(file = file.path(outputdir, "binned-scaled-residuals-3d.RDS"))

## Load the dates
modeldir = "~/repos/flowmixapp/data/01-cytograms/MGL1704/data"
times = readRDS(file = file.path(modeldir, "MGL1704-datobj.RDS"))$time

## ## Saving them once.
## datadir = "~/repos/flowmix/paper-data"
## saveRDS(times, file=file.path(datadir, "binned-residuals-time.RDS"))

## Make 2d plots of a few snapshot in time
iclust = 3
glist = list()
dats_2d = lapply(1:9, function(ii){
  tt = ii * 20
  datobj_2d = collapse_3d_to_2d(
      y = binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]],
      counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]],
      dims = c(1,2)) %>%
    as_tibble() %>%
    add_column(time = times[tt])
}) %>% bind_rows()##.id = "date")
dats_2d = dats_2d %>% mutate(time = as.factor(time))

dats_2d %>% ggplot() + 
  facet_wrap(~time) +
  ## geom_raster(aes(x = diam, y = chl, fill = counts)) +
  geom_raster(aes(x = dim1, y = dim2, fill = counts)) +
  scale_fill_gradientn(colours = c("white", "blue"), guide="colorbar") +
  coord_fixed() +
  ggtitle("After deconvolving")
  ## theme_minimal() 
```

1d plot (diameter) of scaled residuals, by cluster:

## {.tabset}

```{r viz-1d-scaled-residuals, fig.width=10, fig.height=10, results='asis'}
## Get the 1d version.
numclust = 10
TT = 319
dats_by_clust = lapply(1:numclust, function(iclust){ 
  dats = lapply(1:TT, function(tt){
    dt = data.frame(binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]] %>% rbind()) 
    dt = dt %>% add_column( counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]])
    dt = dt %>% setNames(c("diam", "chl", "pe", "counts"))
    if(nrow(dt>0)) dt = dt %>% tibble::add_column(tt)
    return(dt)
  }) %>% bind_rows() %>% as_tibble()
  ## dat_1d = dats %>% group_by(diam, tt) %>% summarize(counts=sum(counts))
  ## return(dat_1d)
  return(dats)
})

## Making a small data table containing time indices and dates.
modeldir = "~/repos/flowmixapp/data/01-cytograms/MGL1704/data"
times = readRDS(file = file.path(modeldir, "MGL1704-datobj.RDS"))$time
timetable = data.frame(tt = 1:TT, times = times)

## ## Temporary: Checking that densities in all clusters sum to 1
## dats_by_clust[[1]]
## binobj %>% length()
## binobj[[1]] %>% names()
## ntlist = sapply(1:319, function(tt){
##   sapply(1:10, function(iclust){
##     binobj[[iclust]]$counts_list[[tt]]
##   }) %>% unlist() %>% sum()
## }) 
## ## End temporary


## ## Begin temporary 2
##   load(file = file.path(outputdir, "residuals-3d.Rdata"), verbose = FALSE)
##   numclust = 10
##   residuals = residuals_obj$residuals_by_cluster
##   countslist = residuals_obj$countslist_by_cluster
##   countslist %>% .[[tt]] %>% .[[iclust]] 
## ## End temporary

## name1 ="diam"
for(name1 in c("diam", "chl", "pe")){## for(labeled in c(TRUE, FALSE)){
  ## cat("###", name1, (if(labeled)"labeled" else ""), fill=TRUE)
  labeled = FALSE
  cat("###", name1, fill=TRUE)
  dat_1d_by_clust = lapply(1:numclust, function(iclust){ 
    dats = dats_by_clust[[iclust]]
    dat_1d = dats %>% group_by(!!sym(name1), tt) %>% summarize(counts = sum(counts))
    ## dat_1d$counts = dat_1d$counts / ntlist[tt]
    dat_1d
  })
  
  dat_1d_by_clust = dat_1d_by_clust %>% bind_rows(.id="iclust")
  dat_1d_by_clust = dat_1d_by_clust %>% mutate(iclust = factor(iclust, levels = 1:10))
  
  if(labeled){
    names = c(pro = c(10,NA), small =NA,  syn = 3,  pico1_ = c(9  ),
                pico2_ = c(1,NA), pico3_ = c( 6), pico4_ = c(NA), croco =NA,
              bead = NA, rest1_ = c(4, 7), rest2_ = c(8), bg = c(5, 2)) %>%
      na.omit() %>%
      sort() %>% names()
  } else {
    names = 1:numclust
  }
  
  ## Add cluster names (doesn't do anything right now)
  dat_1d_by_clust = dat_1d_by_clust %>% mutate(clustername = names[iclust])

  ## Standardize the counts 
  if(FALSE){
    dat_1d_by_clust = dat_1d_by_clust %>% group_by(iclust, tt)  %>% mutate(counts = counts/sum(counts))
  }

  ## Add times
  dat_1d_by_clust = dat_1d_by_clust %>% left_join(timetable)

  
  ## Make plot
  p = dat_1d_by_clust %>% ggplot() + 
    ## facet_wrap(iclust~., nrow=2) +
    facet_wrap(clustername~., nrow = 5) +
    geom_raster(aes(x = times, y = !!sym(name1), fill = counts)) +
    scale_fill_gradientn(colours = c("white", "black", "yellow")) + ##(low="skyblue", high= "black") +
    xlab("Time") +
    ylab("log(cell diameter)") +
    scale_x_datetime(date_breaks = "1 day") +
    theme(axis.text.x = element_text(angle = 25, vjust = 1.0, hjust = 1.0))

  plot(p)
  cat("\n\n")
  ## }
}
```


# Visualize and test 3d residuals

Hereon, we operate with scaled residuals. 

First, obtain the means over time (for each cluster).

```{r form-means} 
means_by_clust = list() 
for(iclust in 1:numclust){
  means = lapply(1:TT, function(tt){
    resid = all_scaled_resids %>% .[[tt]] %>% .$resids %>% .[[iclust]] 
    if(nrow(resid) == 0) return(tibble(diam=NA, chl=NA, pe=NA))
    resid = resid %>% as_tibble() %>% setNames(c("diam", "chl", "pe"))
    count = all_scaled_resids %>% .[[tt]] %>% .$counts %>% .[[iclust]] 
    combined = resid %>% add_column(count)
    avg = combined %>% mutate(prop = count/sum(count)) %>% summarize(diam = sum(diam*prop),
                                                                     chl = sum(chl*prop),
                                                                     pe = sum(pe*prop))
    return(avg)
  })  %>% bind_rows()
  ## print(means %>% apply(2, max, na.rm=TRUE) %>% round(3))
  means_by_clust[[iclust]] = means %>% add_column(times)
} 
```

## One-dimensional means over time

## {.tabset}

```{r plot-1d-means, results='asis', fig.width=15, fig.height=15}
longnames = c(diam = "Diameter (diam)", chl = "Chlorophyll (chl)", pe = "Phycoerythrin (pe)")
for(name1 in c("diam", "chl", "pe")){
  longname1 = longnames[name1]
  cat("###", name1, fill=TRUE) 
  dt = lapply(1:numclust, function(iclust){
    dt0 = tibble(mean = means_by_clust[[iclust]] %>% pull(!!sym(name1)),
                 times = times)
  }) %>% bind_rows(.id = "iclust") %>% mutate(iclust=factor(iclust, levels=1:10))
  p = dt %>% ggplot(aes(x = times, y = mean)) +
    facet_wrap(~iclust, nrow = 5) +
    geom_point(size=rel(.5)) +
    geom_line() +
    ylab(name1)
  p = p + labs(title = paste0("Mean of residuals: ", longname1))
  plot(p)
  cat("\n\n")
}
```

Overlaying the binned residuals (background, in greyscale and yellow) with the
residual means (red).

```{r separate-plots, fig.width=10, fig.height=5}
for(name1 in c("diam", "chl", "pe")){

  longname1 = longnames[name1]

  cat("###", name1, fill=TRUE) 

  ## Get binned data ready
  dat_1d_by_clust = lapply(1:numclust, function(iclust){ 
    dats = dats_by_clust[[iclust]]
    dat_1d = dats %>% group_by(!!sym(name1), tt) %>% summarize(counts=sum(counts))
  })
  
  dat_1d_by_clust = dat_1d_by_clust %>% bind_rows(.id="iclust")
  dat_1d_by_clust = dat_1d_by_clust %>% mutate(iclust = factor(iclust, levels = 1:10))
  dat_1d_by_clust = dat_1d_by_clust %>% mutate(clustername = names[iclust])

  ## Cluster means
  dt = lapply(1:numclust, function(iclust){
    dt0 = tibble(mean = means_by_clust[[iclust]] %>% pull(!!sym(name1)),
                 times = times)
  }) %>% bind_rows(.id = "iclust") %>% mutate(iclust=factor(iclust, levels=1:10))
  
  for(iclust in 1:10){
  
    ## Get the mean ready 
    dt1 = dt %>% filter(iclust == !!(iclust))
    dt1 = dt1 %>% arrange(times) %>% add_column(tt = 1:nrow(dt1))
    
    p = dat_1d_by_clust %>% filter(iclust==!!(iclust)) %>% ggplot() + 
      geom_raster(aes(x=tt, y=!!sym(name1), fill=counts)) +
      scale_fill_gradientn(colours = c("white", "black", "yellow")) 
    p = p + geom_line(aes(x = tt, y = mean), size=rel(.5), data = dt1, size = rel(2), col = 'red') 
    p = p + geom_point(aes(x = tt, y = mean), size=rel(.5), data = dt1, size = rel(.5), col = 'red') 
    p = p + labs(title = paste0("Binned residuals\nCluster ", iclust,
                                " and dimension: ", longname1))
    p = p + xlab("Time index (t)")
    plot(p)
    filename = paste0("1d-", name1, "-clust-", iclust, ".png")
    ggsave(file = file.path(outputdir, filename), width = 10, height = 5)
  }
} 
```

## 2d plots of means over time 

Next, make a 3d plot of (diam, pe, time), over time:

```{r 3d-means-plotly}
name1 = "diam"
name2 = "pe"

## for(iclust in 1:numclust){
htmltools::tagList(lapply(1:numclust, function(iclust) {
  p = plot_ly(
    data = means_by_clust[[iclust]],
    type = 'scatter3d',
    mode = 'lines + markers',
    x = ~ times,
    y = as.formula(paste0('~', name1)),
    z = as.formula(paste0('~', name2)),
    marker = list(size = 3.5, color = ~c, colorscale = 'Greens', cmin = -20, cmax = 50)
  ) %>% layout(scene = list(aspectmode="manual",aspectratio = list( x = 3, y = 1, z = 1))) %>%
    layout(title=paste0("Cluster ", iclust))
  p
})) 
```


## ACF plots

The next set of plots show the autocorrelation function (ACF) plots of the three
measurements (diam, chl, pe) over time, in each cluster.

## {.tabset}

```{r means-acf, fig.width=15, fig.height=5, results='asis'}
for(iclust in 1:numclust){
  cat("### clust", iclust, fill=TRUE)

  ## Make the separate ACF plots
  means = means_by_clust[[iclust]] %>% drop_na()
  par(mfrow=c(1,3))
  acf(means$diam, main = paste("ACF Plot for", "diameter"))
  acf(means$chl , main = paste("ACF Plot for", "chl"))
  acf(means$pe  , main = paste("ACF Plot for", "pe"))
  cat("\n\n")
}
```

## Scatterplots of orig vs lagged means

These plots shows the scatterplots of means in each dimension against their
time-lagged versions.

## {.tabset}

```{r means-lag-plot, fig.width=15, fig.height=15, results='asis'}
for(iclust in 1:numclust){
  cat("### clust", iclust, fill = TRUE)

  ## Make the scatterplots
  means = means_by_clust[[iclust]] %>% drop_na()
  p = plot_lagged(means$diam) + ggtitle(paste0("diam", ", cluster :", iclust))
  print(p)
  name1 = "diam"
  filename = paste0("acf-", name1, "-clust-", iclust, ".png")
  ggsave(file = file.path(outputdir, filename), width = 15, height = 15)

  p = plot_lagged(means$chl) + ggtitle(paste0("chl", ", cluster :", iclust))
  print(p)
  name1 = "chl"
  filename = paste0("acf-", name1, "-clust-", iclust, ".png")
  ggsave(file = file.path(outputdir, filename), width = 15, height = 15)

  p = plot_lagged(means$pe) + ggtitle(paste0("pe", ", cluster :", iclust))
  print(p)
  name1 = "pe"
  filename = paste0("acf-", name1, "-clust-", iclust, ".png")
  ggsave(file = file.path(outputdir, filename), width = 15, height = 15)

  cat('\n\n')
}
```


## Ljung-box test

Next, we apply multivariate Ljung-Box test to the means:


## {.tabset}

```{r print-ljung-means, results = 'asis'}
for(iclust in 1:numclust){
  cat("### clust", iclust, fill=TRUE)
  mq_table(means_by_clust[[iclust]] %>% drop_na() %>% select(diam, chl, pe), 18)
  cat('\n\n')
}
```

# Deviance residuals

Next, we take deviance residuals and plot them over time. 

```{r get-devres}
devres = get_devres(origres, particle_datobj$ylist %>% lapply(as.matrix), particle_datobj$countslist)
ntlist = sapply(particle_datobj$countslist, sum)
devresmat = tibble(times = times,
                   ntlist = ntlist,
                   devres = devres) %>%
  mutate(scaled_devres = devres/ntlist)
```



## Plot deviance residuals

This three-panel plot shows (i) deviance residuals, (ii) $n_t$ over $t=1,..,
319$, and (iii) *scaled* deviance residuals (i.e. divided by $n_t$).

```{r plot-devres, fig.width=10, fig.height=10}
mynames <- c(
  'devres'="Deviance residuals",
  'ntlist'="Biomass by time",
  'scaled_devres'="Scaled deviance residuals")

devresmat %>%  
  pivot_longer(cols = !(times)) %>% 
  ggplot(aes(x=times, y=value)) +
  facet_wrap(~name, scale = "free", nrow=3, labeller=as_labeller(mynames)) +
  geom_point() + 
  geom_line() +
  ylab("") +
  ggtitle("Deviance Residuals")


devresmat %>%  select(times, scaled_devres) %>% 
  ## pivot_longer(cols = !(times)) %>% 
  ggplot(aes(x=times, y=scaled_devres)) +
  ## facet_wrap(~name, scale = "free", nrow=3, labeller=as_labeller(mynames)) +
  geom_point() + 
  geom_line() +
  ylab("") +
  ggtitle("Deviance Residuals")
filename = "devres.pdf"
ggsave(file = file.path(outputdir, filename), width = 10, height = 3.5)
```

This plot shows the scaled deviance residuals' scatterplots against their
time-lagged versions.

```{r, plot-devres2, fig.width = 15, fig.height=15}
devresmat$scaled_devres %>% plot_lagged() + ggtitle(paste0("Scaled deviance residuals"))
```


## Test deviance residuals

Then, we test the deviance residuals (a univariate time series) using
multivariate Ljung-Box.

```{r, test-devres, fig.width=3, fig.height=3, results="asis"}
devresmat$scaled_devres %>% mq_table(17,  caption = "Ljung-box on scaled deviance residuals")
```



## Permutation test


Conduct permutation test of the residuals' means, for each cluster.

```{r permtest, fig.width = 12, fig.height=5}
filename = "perm-test.pdf"
## pdf(file = file.path(outputdir, filename), width = 12, height=5)
par(mfrow = c(2, 5))
for(iclust in 1:10){
  means = means_by_clust[[iclust]]
  test_stat(means)
  permtest(means, 1000, plot = TRUE, title = paste0("Cluster ", iclust))
}
## graphics.off()
```

Also conduct permutation test on the deviance residuals.

```{r permtest-devres, fig.width=5, fig.height=5}
means = devresmat$scaled_devres 
filename = "permtest-devres.pdf"
## pdf(file = file.path(outputdir, filename), width = 5, height=5)
permtest(cbind(means), 1000, plot = TRUE, title = NULL)##"Permutation test\n on Deviance residuals")
## graphics.off()
```

# Using Wasserstein's distance

Instead of cross-correlation :

$$ T(\{\tilde r_{kt}^i\}_{t=1}^T) :=   \max_{\substack{d\in\{1,2,3\},\\ l\in\{1,\cdots, 18\}}} \left|\text{corr}\left(\{e_d^T
\tilde r^i_{kt}\}_{t=1}^{T-l},\;\; \{e_d^T \tilde r^i_{kt} \}_{t=l+1}^T\right) \right|, i = 1, \cdots, 1000. $$

of the (permuted) means:

$$e_d^T \tilde r_{kt}^i,$$


we'll calculate 

$$\{h(e_d^T r_{kt}^j)\}_{t, d, l}$$

across $t, k, $l$.

The thing to measure is

$$W_{t_1,k_1,l_1}^{t_2, k_2, l_2}  = Wass (h(e_{d_1}^T r_{k_1, t_1}^j), h(e_{d_2}^T r_{k_2, t_2}^j)$$


First, a sanity check: does EMD measure correctly?

```{r emd-test-small}
outputdir = "/home/sangwonh/repos/flowmix/paper-code/aoas-revision/3d-residuals"
maindir = "/home/sangwonh/repos/flowmix/paper-code/aoas-revision"
library(ggplot2)
library(tidyverse)
la("flowmix")
name1 = "dim1"
name2 = "dim2"
iclust = 3
binobj = readRDS(file = file.path(outputdir, "binned-scaled-residuals-3d.RDS"))
plist = lapply(c(100, 101, 102, 103, 104, 200), function(tt){
  collapse_3d_to_2d(
      y = binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]],
      counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]],
      dims = c(1,2)) %>%
    as_tibble() %>%
    ggplot() +
    geom_raster(aes(x = !!sym(name1), y = !!sym(name2), fill = counts)) +
    scale_fill_gradientn(colours = c("white", "blue"), guide="colorbar")  +
    coord_fixed() +
    ylim(c(-3,3)) +
    xlim(c(-3,3)) +
    ggtitle(paste0("time = ", tt))
})
do.call(gridExtra::grid.arrange, c(plist, ncol = 3))

sixdat = lapply(c(100, 101, 102, 103, 104, 200), function(tt){
  dat = data.frame(binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]],
                   counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]])
})

## Collapse 1o 1d
iclust = 3
ylist_small = lapply(c(100, 101, 102, 103, 104, 200), function(tt){
  dat = data.frame(binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]]) %>% as.matrix()
})
countslist_small = lapply(c(100, 101, 102, 103, 104, 200), function(tt){
  counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]] %>% as.numeric()
})

## What the problem
dat1d = collapse_3d_to_1d(ylist_small, countslist_small, idim = 2)
ylist_1d = dat1d$ylist
countslist_1d = dat1d$countslist

dmat = matrix(NA, nrow = 6, ncol = 6)
for(ii in 1:6){
  for(jj in 1:6){
    dmat[ii, jj] = emd(ylist_1d[[ii]] %>% cbind(),
                       countslist_1d[[ii]],
                       ylist_1d[[jj]] %>% cbind(),
                       countslist_1d[[jj]])

  }
}
colnames(dmat) = rownames(dmat) =  c(100, 101, 102, 103, 104, 200)
drawmat_precise(dmat)
```

It appears so!


```{r emd-test-large, eval=FALSE}
## Okay so the next thing to try is making a large matrix, and matching the rows & columns?
all_y = binobj %>% .[[iclust]] %>% .$ybin_list %>% do.call(rbind, .) %>% as_tibble()
colnames(all_y) = c("diam", "chl", "pe")
all_y %>% pull(diam) %>% unique() %>% sort() %>% length()
all_y %>% pull(chl) %>% unique() %>% sort() %>% length()
all_y %>% pull(pe) %>% unique() %>% sort() %>% length()

## Rows are "from", columns are "to"
TT = 319
## for(iclust in
mclapply(c(3,1,2,4:10), function(iclust){
  for(idim in 1:3){
    if(file.exists(file = file.path(outputdir, "distmats", paste0(idim, "-", iclust, ".RDS")))) next
    ylist = lapply(1:TT, function(tt){
      dat = data.frame(binobj %>% .[[iclust]] %>% .$ybin_list %>% .[[tt]]) %>% as.matrix()  })
    countslist = lapply(1:TT, function(tt){
      counts = binobj %>% .[[iclust]] %>% .$counts_list %>% .[[tt]] %>% as.numeric()  })
    dat1d = collapse_3d_to_1d(ylist, countslist, idim = idim)
    cat(fill = TRUE)

    ## There is one that is both NA; what to do?
    ylist_1d = dat1d$ylist
    countslist_1d = dat1d$countslist

    ## Calculate the distance
    dists = matrix(NA, nrow = TT, ncol = TT)
    for(tt1 in 1:TT){
      for(tt2 in 1:TT){
        if(abs(tt1 - tt2) >= 100) next

        if(any(is.na(ylist_1d[[tt1]])) | any(is.na(ylist_1d[[tt2]]))) next
        if(any(is.na(countslist_1d[[tt1]])) | any(is.na(countslist_1d[[tt2]]))) next
         dists[tt1, tt2] =  emd(ylist_1d[[tt1]] %>% cbind(),
              countslist_1d[[tt1]],
              ylist_1d[[tt2]] %>% cbind(),
              countslist_1d[[tt2]])
      }
    }
    colnames(dists) = rownames(dists) =  1:TT
    saveRDS(dists, file = file.path(outputdir, "distmats", paste0(idim, "-", iclust, ".RDS")))
    cat(fill = TRUE)
    cat("saved to", file.path(outputdir, "distmats", paste0(idim, "-", iclust, ".RDS")))
  }
  return(NULL)
}, mc.cores = 2)


## Drawing one matrix at a time
for(idim in 1:3){
  for(iclust in 1:10){
    iclust=1
    dists = readRDS(file = file.path(outputdir, "distmats", paste0(idim, "-", iclust, ".RDS")))
    drawmat_precise(dists)
  }
}
```

Now, let's make a plot of cross-correlations that use Wasserstein distance
instead of Pearson correlation.


```{r main-emd-plot, fig.width = 10, fig.height = 14}
## Okay, so from this, we can easily calculate the cross-correlations
dats = list()
for(idim in 1:3){
  for(iclust in 1:10){

    ## Load distance matrix
    dists = readRDS(file = file.path(outputdir, "distmats", paste0(idim, "-", iclust, ".RDS")))

    ## Get all means
    allstats = sapply(1:100, function(numlag){
      onestat = dists[row(dists) == (col(dists) - numlag)] %>% mean(na.rm=TRUE)
      return(onestat)
    })

    ## Save as a long 
    onedat = data.frame(lag = 1:100, dist = allstats, idim = idim, iclust = iclust) %>% as_tibble()
    onename = paste0(idim, "-", iclust)
    dats[[onename]] = onedat
  }
}

## String together by row:
long = do.call(rbind, dats)
long = long %>%
  mutate(idim = replace(idim, idim == 1, "diam")) %>%
  mutate(idim = replace(idim, idim == 2, "chl")) %>%
  mutate(idim = replace(idim, idim == 3, "pe"))
long = long %>% mutate(idim = factor(idim, levels = c("diam", "chl", "pe")))

p = long %>%
  ggplot() +
  facet_grid(iclust ~ idim) +
  geom_point(aes(x = lag, y = dist), size = rel(0.3)) +
  geom_line(aes(x = lag, y = dist), alpha=.5) +
  ylab("Mean Wasserstain Distance") +
  xlab("Number of Time Lags") +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Cluster number", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Cytogram dimension", breaks = NULL, labels = NULL))
plot(p)
ggsave(file = file.path(outputdir, "mean-wasserstein.pdf"), width = 10, height = 14)
```


<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height:0;"></div>



```{r}
knitr::knit_exit() 
```

The remainder is a graveyard.

# Combine the residuals

```{r, eval = FALSE}
all_scaled_resids %>% length()

residlist = list()
for(tt in 1:TT){
  dt_combined = lapply(1:numclust, function(iclust){
    rr = all_scaled_resids %>% .[[tt]] %>% .$resid %>% .[[iclust]]
    cc = all_scaled_resids %>% .[[tt]] %>% .$counts %>% .[[iclust]]
    dt = data.frame(rr) %>% add_column(counts=cc) %>% as_tibble()
    colnames(dt) = c("diam", "chl", "pe", "counts")
    dt
  }) %>% bind_rows()
  residlist[[tt]] = dt_combined
}

dat_grid = make_grid(residlist, 100)[1:3] %>% setNames(c("diam", "chl", "pe"))
## ranges = lapply(residlist,
##                 function(a){ apply(a, 2, range) %>% as_tibble()}) %>%
##   bind_rows() %>%
##   apply(2, range)
## ranges

diam = residlist[[1]] %>% pull(diam) 

residlist[[2]] %>% pull(diam) %>% hist()

chl = residlist[[2]] %>% pull(chl) 
hist(chl)

chl[chl>5]
counts[chl>5]

counts = residlist[[1]] %>% pull(counts)
max(chl)
diam[diam< -5]

## tt = 1
ylist = list()
countslist = list()
for(tt in 1:TT){
  print(tt)
  dt_combined = residlist[[tt]]
  one_bin = bin_one_cytogram(dt_combined[,c("diam","chl", "pe")],
                             qc = dt_combined[,"counts", drop=TRUE],
                             manual.grid = dat_grid)
  ylist[[tt]] = one_bin$ybin
  countslist[[tt]] = one_bin$counts
}

collapse_3d_to_1d(ylist, countslist, idim = 1)

```

